### 2018-03-30 kanwal-analysis/BDA-analysis.Rmd
R analysis  -- this helps because we can run the model from R and not worry about cutting off the outputs (as we have running from command line)

---
title: "Kanwal (2017) BDA Notebook"
author: "Ben"
date: "3/30/2018"
output: html_document
---

```{r packages, message=FALSE, warning=FALSE}
library(dplyr)
library(ggplot2)
library(ggthemes)
library(rwebppl)
library(tidyr)
```

```{r echo=FALSE}
# getModelFile()
# ==============
# Return a model string from a .wppl file
#
# Parameters
# ----------
# modelFile: str
#  File path str
# 
# Returns
# -------
# str
#    Model str.
#
getModelFile <- function(modelFile) {
  readChar(modelFile, file.info(modelFile)$size)
}
```

## BDA data
```{r generative-model}
fPath <- "../kanwal-generative-model-20180328.wppl"
m <- getModelFile(fPath)
```

Get posterior predictives...
```{r run-webppl}
dfSynthetic <- rwebppl::webppl(m)
```

## Empirical Kanwal (2017) data
```{r get-empirical-data}
fEmpiricalPath <- "../kanwal-preprocessed.csv" 
dfEmpirical <- read.csv(fEmpiricalPath)
```

```{r prepocessing-synthetic-data-names}
# Minor preprocessing
dfSynthetic <- dfSynthetic %>%
  rename(display_type=trialType) %>%
  mutate(label_type=ifelse(utterance=='zop', 'short', 'long'))
```

## Preprocessing
```{r prepocessing-synthetic-data-counts}
dfSyntheticCnts <-  dfSynthetic %>%
  group_by(IP, condition, display_type, label_type) %>%
  summarise(cnt=n())
```

Fill missing vals synthetic data (note I've alread carried this out in `analysis.ipynb` for empirical data)
```{r prepocessing-synthetic-data-fill-missing-vals}
display_types <- c('frequent', 'infrequent')
label_types <- c('short', 'long')

# New data frame for merging missing counts
dfPlaceholder <- data.frame(expand.grid(IP=unique(dfSynthetic$IP),
                                        condition=unique(dfSynthetic$condition),
                                        display_type=display_types,
                                        label_type=label_types)) %>%
  mutate(
    IP=as.character(IP),
    condition=as.character(condition),
    display_type=as.character(display_type),
    label_type=as.character(label_type),
    cnt=as.integer(0))

# First combine and take counts
dfJoin1 <- full_join(dfSyntheticCnts, dfPlaceholder, by=c('IP', 'condition', 'display_type', 'label_type', 'cnt')) %>%
  arrange(IP, condition, display_type) %>%
  group_by(IP, condition, display_type, label_type) %>%
  summarise(cnt=sum(cnt))

# We want to filter based on whether the group (IP, condition) ash *any* counts, so get these IDs
dfJoin2 <- full_join(dfSyntheticCnts, dfPlaceholder, by=c('IP', 'condition', 'display_type', 'label_type', 'cnt')) %>%
  arrange(IP, condition, display_type) %>%
  group_by(IP, condition, display_type, label_type) %>%
  summarise(cnt=sum(cnt)) %>%
  group_by(IP, condition) %>%
  summarise(groupTotal=sum(cnt))

dfSyntheticFinal <- full_join(dfJoin1, dfJoin2, by=c('IP', 'condition')) %>%
  filter(groupTotal == 32) %>%
  mutate(condition_total=ifelse(display_type=='frequent', 24, 8)) %>%
  group_by(IP, condition, display_type) %>%
  mutate(prop=cnt/condition_total) %>%
  select(IP, condition, display_type, label_type, prop) %>%
  filter(label_type=='short')
  # spread(display_type, prop)
```

Join empirical data and synthetic data
```{r preprocessing-join-empirical-synthetic-data}
dfSyntheticFinal <- dfSyntheticFinal %>%
  select(IP, condition, display_type, prop)

dfEmpiricalProcessed <- dfEmpirical %>%
  mutate(IP=as.character(IP),
         condition=as.character(condition)) %>%
  gather(display_type, prop, c(frequent, infrequent)) %>%
  select(IP, condition, display_type, prop)

dfEmpSynth <- left_join(dfSyntheticFinal, dfEmpiricalProcessed, by=c("IP", "condition", "display_type")) %>%
  gather(data_type, prop, c(prop.y, prop.x)) %>%
  select(IP, condition, display_type, data_type, prop) %>%
  mutate(data_type=ifelse(data_type=='prop.x', 'synthetic', 'empirical'))
```

### Evaluation

How do model predictions look compared to empirical data?
```{r plot-synthetic-vs-empirical}
dfEmpSynth %>%
  spread(display_type, prop) %>%
  ggplot(aes(x=frequent, y=infrequent, col=data_type, group=IP)) +
    geom_point(alpha=0.7, size=2) +
    geom_line(alpha=0.7, col='grey', lty=2) +
    theme_few() +
    facet_grid(~condition)
```

Correlation
```{r correlation-synthetic-vs-empirical}
dfEmpSynthCor <- dfEmpSynth  %>%
  spread(data_type, prop)

paste("R^2:", round(cor(dfEmpSynthCor$empirical, dfEmpSynthCor$synthetic)^2, 3))
```

```{r plot-correlation}
dfEmpSynth %>%
  spread(data_type, prop) %>%
  ggplot(aes(x=synthetic, y=empirical)) +
    geom_jitter(aes(col=condition), alpha=0.5, width=0.01, height=0.01, size=2) +
    geom_smooth(method='lm') +
    ggtitle("Model vs Empirical predictions Kanwal (2017)") +
    theme_few() +
    theme(plot.title = element_text(hjust = 0.5))
```

### 2018-03-31 kanwal-analysis/BDA-analysis.Rmd
updated analysis on hueristic modelable subset

@@
 
-## BDA data
+# BDA - Modeling participant data
+
+Note: this analysis includes 85% of participants as I'm having trouble modeling what appears to be fairly random behaviors... See `analysis.ipynb` for more information.
+```{r getModelableData}
+# Hueristic modelable data
+f_modelable_data <- "../modelable-data-20180330.csv"
+dfModelable <- read.csv(f_modelable_data)
+
+# All data
+f_all_data <- "../kanwal-preprocessed.csv"
+dfAll <- read.csv(f_all_data) %>%
+  mutate(frequent=round(frequent, 2),
+         infrequent=round(infrequent, 2))
+```
+
+```{r setup-bda-model}
+f_bda_model <- "../kanwal-BDA-20180330.wppl"
+model_bda <- getModelFile(f_bda_model)
+bda_data <- head(dfModelable)
+nrow(dfModelable)
+
+x1 <- '115.98.55.114'
+x2 <- '117.217.210.9'
+# but worked on second time
+dfAll2 <- dfAll %>% filter(IP != x2,
+                           condition !='neither')
+```
+
+Note: slow running.
+```{r run-bda-model}
+# pt <- proc.time()
+# df_bda <- rwebppl::webppl(model_bda, data=dfModelable, data_var='rData')
+# et <- proc.time() - pt
+# write.csv(df_bda, "bda-tuned-params-20180303.csv")
+```
+
+```{r optional-read-cached-data}
+# If we haven't re-run the BDA predictives then load cached version.
+if (!exists('df_bda')) {
+  df_bda <- read.csv("bda-tuned-params-20180303.csv")
+}
+```
+
+## BDA predictive
 ```{r generative-model}
@@
 ```{r run-webppl}
-dfSynthetic <- rwebppl::webppl(m)
+dfSynthetic <- rwebppl::webppl(m, data=df_bda, data_var='rData')
 ```
@@
   spread(display_type, prop) %>%
-  ggplot(aes(x=frequent, y=infrequent, col=data_type, group=IP)) +
-    geom_point(alpha=0.7, size=2) +
+  ggplot(aes(x=frequent, y=infrequent, col=condition, group=IP)) +
+    geom_point(aes(shape=data_type), alpha=0.5, size=2.5) +
     geom_line(alpha=0.7, col='grey', lty=2) +
@@
 ```
+
+### Restricting analysis to rational scenarios...
+```{r correlation-synthetic-vs-empirical-conditions}
+dfEmpSynthCor <- dfEmpSynth  %>%
+  filter(condition != 'neither') %>%
+  spread(data_type, prop)
+
+paste("R^2:", round(cor(dfEmpSynthCor$empirical, dfEmpSynthCor$synthetic)^2, 3))
+```
+
+```{r plot-correlation-conditions}
+dfEmpSynth %>%
+  filter(condition != 'neither') %>%
+  spread(data_type, prop) %>%
+  ggplot(aes(x=synthetic, y=empirical)) +
+    geom_jitter(aes(col=condition), alpha=0.5, width=0.01, height=0.01, size=2) +
+    geom_smooth(method='lm') +
+    ggtitle("Model vs Empirical predictions Kanwal (2017)") +
+    theme_few() +
+    theme(plot.title = element_text(hjust = 0.5))
+```

### 2018-03-31 kanwal-analysis/BDA-analysis.Rmd
added baseline mode comparison. baseline simply chooses the optimal strategy in all of the contidions (ignoring "neither" condition).

@@
 
-x1 <- '115.98.55.114'
-x2 <- '117.217.210.9'
+problemIPs <- c('70.198.5.150', 
+                '117.217.210.9', 
+                '65.183.140.246', 
+                '24.253.233.194', 
+                '24.20.229.200', 
+                '108.196.136.202',
+                '122.161.26.31')
+# 24.74.13.244 only after 10000 attempts
+# 50.153.162.232
+# 
 # but worked on second time
-dfAll2 <- dfAll %>% filter(IP != x2,
-                           condition !='neither')
+dfAll2 <- dfAll %>% filter(!(IP %in% problemIPs))
 ```
@@
 # pt <- proc.time()
-# df_bda <- rwebppl::webppl(model_bda, data=dfModelable, data_var='rData')
+# df_bda <- rwebppl::webppl(model_bda, data=dfAll2, data_var='rData')
 # et <- proc.time() - pt
@@
 ```
+
+### Baseline comparison  - naive model
+
+The naive model employs the optimal strategy for communication in each of the three conditions.
+* `combined`: always use `short` form for frequent object, always use `long` form for infrequent object
+* `accuracy`: always use `long` form
+* `time`: always use `short` form
+```{r}
+dfEmpSynthAdd <- dfEmpSynth %>%
+  filter(data_type=='empirical') %>%
+  mutate(baseline=ifelse((condition=='combined' & display_type=='frequent') | 
+                               (condition=='accuracy' & display_type=='infrequent') |
+                               (condition=='time'), 1.0, 0.0)) %>%
+  select(IP, condition, display_type, baseline) %>%
+  rename(prop=baseline) %>%
+  mutate(data_type='baseline')
+
+dfEmpSynthBaseline <- rbind(dfEmpSynth, dfEmpSynthAdd)
+```
+
+```{r plot-model-comparison-data}
+dfEmpSynthBaseline %>%
+  filter(condition != 'neither') %>%
+  spread(display_type, prop) %>%
+  ggplot() +
+    geom_jitter(aes(x=frequent, y=infrequent, col=condition), alpha=0.5, width=0.01, height=0.01, size=2) +
+    ggtitle("Model vs Empirical predictions Kanwal (2017)") +
+    theme_few() +
+    theme(plot.title = element_text(hjust = 0.5)) +
+    facet_grid(~data_type)
+```
+
+```{r plot-model-comparison-corr}
+dfEmpSynthBaseline %>%
+  filter(condition != 'neither') %>%
+  spread(data_type, prop) %>%
+  gather(model, prop, c(baseline, synthetic)) %>%
+  ggplot(aes(x=prop, y=empirical)) +
+    geom_point(alpha=0.5, size=2) +
+    geom_smooth(method='lm') +
+    ggtitle("Model vs Empirical predictions Kanwal (2017)") +
+    theme_few() +
+    theme(plot.title = element_text(hjust = 0.5)) +
+    facet_grid(~model)
+```
+
+```{r correlation-baseline}
+dfBaselineCor <- dfEmpSynthBaseline %>%
+  filter(condition != 'neither') %>%
+  spread(data_type, prop)
+
+cor(dfBaselineCor$baseline,
+    dfBaselineCor$empirical)^2
+
+```

### 2018-06-13 analysis/BDA-analysis.Rmd
re-organize dirs

--- 2018-03-31 kanwal-analysis/BDA-analysis.Rmd
@@
 
-```{r packages, message=FALSE, warning=FALSE}
+[Kanwal (2017)](https://www.ncbi.nlm.nih.gov/pubmed/28494263) present a test of Zipf's `Law of Abbreviation` -- that a universal design feature of human language emerges as a result of individuals optimising form-meaning mappings under competing pressures to communicate accurately (`expressivity`), but also efficiently (roughly, `learnabilty`).
+Their study uses miniature artificial language learning paradigm.
+They show that language users optimise form-mean mapping only when pressures for accuracy and effficency *both* operate during the communicative task.
+
+In this analysis we explore the degree to which the observed behavior can be explained by more general purpose social reasoning mechanisms using a computational pragmatics model.
+We show that our model accounts for a large portion of the variance ($r^2=0.97$) across multiple conditions (both optimal and sub-optimal behavior).
+
+While one interpretation of this finding frames the results as entirely pragmatics-driven. Instead we argue that the Zipfian result is a bi-product of pragmatic mechanims.
+
+Summary of paper:
+
+Four conditions:
+
+  * `combination`: both communicative (expressivity) and production cost (efficiency) pressures.
+
+  * `accuracy`: only communicative pressure (no production cost).
+
+  * `timing`: no communicative pressures (only production costs).
+
+  * `neither`: not communicative pressures or productino cost. Note this scenario is not very interesting to us as it does not resemble any aspects of communication.
+
+  ![Kanwal (2017), fig3](img/kanwal-2017-fig3.png)
+
+```{r packages, message=FALSE, warning=FALSE, echo=FALSE}
 library(dplyr)
@@
 
-# BDA - Modeling participant data
+# BDA - Modeling Zipf's Law of Abbreviation via Kanwal (2017)
+
+Here we model participant behavior as pragmatic reasoning instantiated in the Rational speech-act family of models.
+We infer four hyperparmeters which encode different participant strategies (optimal in different conditions).
+
+### BDA generative model
+
+* `semantics`: $Unif(["natural", "none"])$ where
+    * $natural: [[zop]] \in \{zopekil, zopudon\}, [[zopudon]] \in \{zopudon\}, [[zop]] \in \{zopekil\} \in \{zopekil\}$
+    * $none: [[x]] \in \{zopekil, zopudon\}$
+
+* `costs`: $Unif(["present", "not-present", "reversed"])$ where
+    * $present: "zop":0.3, "zopudon":0.7, "zopekil":0.7$
+    * $present: cost(zop) = cost(zopekil) = cost(zopudon) = 0$
+    * $reversed: "zop":0.7, "zopudon":0.3, "zopekil":0.3$
+
+* `recursion-level`: $Categorical([1, 2])$
+
+* `needProbabilities`: $Categorical([present, not present, reversed])$
+    * `present`: "zopudon":24,"zopekil":8,  
+    * `not-present`: "zopudon":1,"zopekil":1,
+    * `reversed`: "zopudon":8,"zopekil":24,
+
+### Strategy mapping to hyperparams
+
+Three conditions of interest:
+
+  * `combination`:  `semantics`='natural', `costs`='present', `needProbabilities`='present'
+
+  * `accuracy`:  `semantics`='natural', `costs`='not-present'
+
+  * `timing`:  `semantics`='not-present', `costs`='present'
+
+## Analysis
 

### 2018-06-17 analysis/data-explore.Rmd
plotting functionality for individual IP behavior -- gotta figure out what the crazy people are doing.

---
title: "data-explore"
author: "Ben"
date: "6/16/2018"
output: html_document
---

```{r libs}
library(ggthemes)
library(tidyverse)
```

```{r options}
options(stringsAsFactors = FALSE)
```

```{r data}
df <- read.csv("../data/df_agg.csv")
```

```{r add-match-label}
df <- df %>%
  mutate(match=label_type_num==display_type_num)
```

```{r get-conditions}
conditions <- unique(df$condition)
```

```{r plotting-helpers}
plot_IP_behavior_match <- function(cond, df, save_file=TRUE) {
  test_img <- df %>%
    filter(condition == cond) %>%
    ggplot(aes(x=as.factor(trial), y=match, col=display_type, group=1)) +
      geom_point(aes(shape=label_type), size=3) +
      geom_line() +
      theme_few() +
      facet_wrap(~IP, ncol=5) +
      ggtitle(paste("Condition:", cond)) +
      theme(text = element_text(size=12),
            axis.text.x = element_text(angle=90, hjust=1, size=6),
            plot.title = element_text(hjust = 0.5))
  if (save_file) ggsave(paste0("img/test-img-", cond, "-20180616.png"), test_img, width=30, height=30)
}

plot_IP_behavior_label_type <- function(cond, df, save_file=TRUE) {
  test_img <- df %>%
    filter(condition == cond) %>%
    ggplot(aes(x=as.factor(trial), y=label_type, col=display_type, group=1)) +
      geom_point(size=3, alpha=0.5) +
      geom_line() +
      theme_few() +
      facet_wrap(~IP+condition, ncol=5) +
      ggtitle(paste("Condition:", cond)) +
      theme(text = element_text(size=24),
            axis.text.x = element_text(angle=90, hjust=1, size=6),
            plot.title = element_text(hjust = 0.5))
  if (save_file) ggsave(paste0("img/label-type-", cond, "-20180616.png"), test_img, width=30, height=30)
}
```

```{r plots}
for (i in 1:length(conditions)) {
  cond <- conditions[i]
  cat(paste("\nCondition:", cond))
  plot_IP_behavior_match(cond, df)
  plot_IP_behavior_label_type(cond, df)
}
```

```{r bad-ips}
plot_IP_behavior_label_type("accuracy", df)
```

### 2019-01-07 bda/analysis/BDA-analysis.Rmd
update directory structure. add paper outline.

--- 2018-06-13 analysis/BDA-analysis.Rmd
@@
 # Hueristic modelable data
-f_modelable_data <- "../modelable-data-20180330.csv"
+f_modelable_data <- "../data/modelable-data-20180330.csv"
 dfModelable <- read.csv(f_modelable_data)
@@
 # All data
-f_all_data <- "../kanwal-preprocessed.csv"
+f_all_data <- "../data/kanwal-preprocessed.csv"
 dfAll <- read.csv(f_all_data) %>%
@@
 ```{r setup-bda-model}
-f_bda_model <- "../kanwal-BDA-20180330.wppl"
+f_bda_model <- "../models/kanwal-BDA-20180330.wppl"
 model_bda <- getModelFile(f_bda_model)
@@
 problemIPs <- c('70.198.5.150', 
-                '117.217.210.9', 
+                '117.217.210.9',   # this ones the killer
                 '65.183.140.246', 
@@
                 '24.20.229.200', 
-                '108.196.136.202',
+                '108.196.136.202',  # this ones not bad
                 '122.161.26.31')
@@
 # but worked on second time
-dfAll2 <- dfAll %>% filter(!(IP %in% problemIPs))
+# dfAll2 <- dfAll %>% filter(!(IP %in% problemIPs))
+dfAll2 <- dfAll %>% filter(IP %in% problemIPs)
 ```
@@
 ```{r run-bda-model}
-# pt <- proc.time()
-# df_bda <- rwebppl::webppl(model_bda, data=dfAll2, data_var='rData')
-# et <- proc.time() - pt
+pt <- proc.time()
+df_bda <- rwebppl::webppl(model_bda, data=dfAll2, data_var='rData')
+et <- proc.time() - pt
 # write.csv(df_bda, "bda-tuned-params-20180303.csv")

### 2019-01-07 bda/analysis/data-explore.Rmd
update directory structure. add paper outline.

--- 2018-06-17 analysis/data-explore.Rmd
@@
 ```{r libs}
+rm(list=ls())
 library(ggthemes)
+library(jsonlite)
 library(tidyverse)
@@
 
+# Reproduce Fig 5.
+
+```{r}
+f_all_data <- "../data/df_agg.csv"
+dfAll <- read.csv(f_all_data)
+```
+
+```{r}
+dfAll %>%
+  filter(condition=='combined') %>%
+  select(IP, condition, label, conditional_trial_num, label_type, display_type) %>%
+  group_by(conditional_trial_num, display_type)
+
+```
+
+```{r}
+dfAll %>%
+  filter(condition=='combined') %>%
+  select(IP, condition, label, conditional_trial_num, label_type, display_type) %>%
+  mutate(wordLength=nchar(label)) %>%
+  group_by(conditional_trial_num, display_type) %>%
+  summarise(n=n(),
+            avgLen=mean(wordLength),
+            ci_high=avgLen+1.96*sd(wordLength)/sqrt(n),
+            ci_low=avgLen-1.96*sd(wordLength)/sqrt(n)) %>%
+  ggplot(aes(x=conditional_trial_num, y=avgLen, col=display_type)) +
+    geom_point() +
+    geom_line() +
+    geom_errorbar(aes(ymax=ci_high, ymin=ci_low), col='grey', alpha=0.7, width=0.5) +
+    ylim(3, 7) +
+    theme_few() +
+    ggtitle("Reproducing Fig 5.")
+```
+
+```{r}
+dfPropData <- dfAll %>%
+  filter(condition=='combined') %>%
+  select(IP, condition, label, conditional_trial_num, label_type, display_type) %>%
+  mutate(isShortLabel=label_type=='short') %>%
+  group_by(conditional_trial_num, display_type) %>%
+  summarise(n=n(),
+            propShort=mean(isShortLabel),
+            ci_high=propShort+1.96*sqrt(propShort*(1-propShort)/n),
+            ci_low=propShort-1.96*sqrt(propShort*(1-propShort)/n))
+
+dfPropData %>%
+  ggplot(aes(x=conditional_trial_num, y=propShort, col=display_type)) +
+    geom_errorbar(aes(ymax=ci_high, ymin=ci_low)) +
+    geom_line() +
+    ylim(0, 1) +
+    theme_few() +
+    ggtitle("Proportion of times using the short form over repetition.")
+
+# write.csv(dfPropData, "../data/temporalDynamicsData.csv")
+```
+
 ```{r data}
-df <- read.csv("../data/df_agg.csv")
+df <- read.csv("../data/d _agg.csv")
 ```

### 2019-01-07 bda/analysis/fit_model.Rmd
update directory structure. add paper outline.

---
title: "run_model_fitting"
author: "Ben"
date: "11/16/2018"
output: html_document
---

```{r libraries}
library(ggplot2)
library(ggthemes)
library(rwebppl)
library(tidyverse)
```

```{r}
# getModelFile()
# ==============
# Return a model string from a .wppl file
#
# Parameters
# ----------
# modelFile: str
#  File path str
# 
# Returns
# -------
# str
#    Model str.
#
getModelFile <- function(modelFile) {
  readChar(modelFile, file.info(modelFile)$size)
}
```

```{r}
# All data
f_all_data <- "../data/counts_all.csv"
dfAll <- read.csv(f_all_data)
cat("n=", length(dfAll$IP))
```

```{r}
f_bda_model <- "../models/bda.wppl"
model_bda <- getModelFile(f_bda_model)
```

```{r}
problemIPs <- c('142.196.192.44', 
                '173.242.228.167',
                '104.57.161.213',
                '104.11.62.66',
                '184.155.134.46',
                '67.168.183.190',
                '108.210.65.252')
```

```{r}
# pt <- proc.time()
# df_bda <- rwebppl::webppl(model_bda, data=dfAll, data_var='rData', cores=4)
# et <- proc.time() - pt
# cat(paste("Run time: ", et))

df_predictives <- read.csv("../analysis/df_predictives_20181119.csv")
```

```{r merge bda and raw data}
df_predictives <- merge(dfAll, df_predictives)
# write_csv(df_predictives, 'df_predictives_20181119.csv')
```

Check out Model / Human `r^2`
```{r}
df_frequent <- df_predictives %>% 
  select(IP, condition, alpha, depth, inferredCondition, frequent_cnt, posteriorPredictiveFrequent) %>%
  mutate(type='frequent') %>%
  rename(actualCnt=frequent_cnt,
         estCnt=posteriorPredictiveFrequent)

df_infrequent <- df_predictives %>% 
  select(IP, condition, alpha, depth, inferredCondition, infrequent_cnt, posteriorPredictiveInfrequent) %>%
  mutate(type='infrequent') %>%
  rename(actualCnt=infrequent_cnt,
         estCnt=posteriorPredictiveInfrequent)

df_predictives2 <- rbind(df_frequent, df_infrequent)
cor(df_predictives2$actualCnt, df_predictives2$estCnt)^2

```

```{r}
df_predictives2 %>%
  # mutate(actualFrequentProps=frequent_cnt / frequent_n,
  #        actualInrequentProps=infrequent_cnt / infrequent_n,
  #        estFrequentProps=posteriorPr5edictiveFrequent / frequent_n,
  #        estInfrequentProps=posteriorPredictiveInfrequent / infrequent_n) %>%
  # select(IP, condition, inferredCondition, actualFrequentProps, estFrequentProps, actualInrequentProps, estInfrequentProps) %>%
  # gather(type, value, c(posteriorPredictiveFrequent, posteriorPredictiveInfrequent))
  ggplot(aes(x=actualCnt, y=estCnt, col=condition), alpha=0.4) +
    geom_smooth(method='lm') +
    geom_point(aes(fill=condition), alpha=0.6, size=3, colour="white", pch=21) +
    # geom_point(aes(x=actualInrequentProps, y=estInfrequentProps, col=condition), alpha=0.4) +
    facet_wrap(~condition, scales="free") +
    ylab('Posterior predictive \n counts of short form') +
    xlab('Human data \n counts of short form') +
    theme_few()

ggsave('../../paper/figs/post-predictives-plot.png', width=6, height=3)
```

Compare to "optimal baseline" optimal in that `time` condition always pick short, `accuracy` always long, etc.
```{r}
df_predictives2 %>%
  mutate(baseline=ifelse(condition=='combined' & type=='frequent', 24,
                         ifelse(condition=='combined' & type=='infrequenty', 8,
                                ifelse(condition=='accuracy' & type=='frequent', 0,
                                       ifelse(condition=='accuracy' & type=='infrequent', 24,
                                              ifelse(condition=='time' & type=='frequent', 48, 16)))))) %>%
  ggplot(aes(x=actualCnt, y=baseline, col=condition), alpha=0.4) +
    geom_point(aes(fill=condition), alpha=0.6, size=3, colour="white", pch=21) +
    # geom_point(aes(x=actualInrequentProps, y=estInfrequentProps, col=condition), alpha=0.4) +
    facet_wrap(~condition, scales="free") +
    ylab('Baseline model \n counts of short form') +
    xlab('Human data \n counts of short form') +
    theme_few()

df_predictives2 %>%
  mutate(baseline=ifelse(condition=='combined' & type=='frequent', 24,
                         ifelse(condition=='combined' & type=='infrequenty', 8,
                                ifelse(condition=='accuracy' & type=='frequent', 0,
                                       ifelse(condition=='accuracy' & type=='infrequent', 24,
                                              ifelse(condition=='time' & type=='frequent', 48, 16)))))) %>%
  group_by(condition) %>%
  summarise(r2 = cor(actualCnt, estCnt)^2)

df_predictives2 %>%
  mutate(baseline=ifelse(condition=='combined' & type=='frequent', 24,
                         ifelse(condition=='combined' & type=='infrequenty', 8,
                                ifelse(condition=='accuracy' & type=='frequent', 0,
                                       ifelse(condition=='accuracy' & type=='infrequent', 24,
                                              ifelse(condition=='time' & type=='frequent', 48, 16)))))) %>%
  summarise(r2 = cor(actualCnt, baseline))
```

How well do we infer the strategies?
```{r}
mean(df_predictives$inferredCondition == df_predictives$actualCondition)
```

Output inferred subject-level params
```{r}
df_predictives %>%
  select(IP, alpha, depth, inferredCondition) %>%
  toJSON() %>%
  write_lines('subjectInferredParams.json')
```

# Try to model time dynamics

```{r}
f_temporal_bda_model <- "../models/bdaFig5.wppl"
model_temporal_bda <- getModelFile(f_temporal_bda_model)
df_trajectories <- rwebppl::webppl(model_temporal_bda)
```

```{r}

df2 <- data.frame()
for (i in seq(1, length(df_trajectories))) {
  df2 <- rbind(df2, df_trajectories[[i]])
}

```
```{r}
numIPs <- length(unique(df2$IP))
df2$trialNum <- rep(seq(1, 31), numIPs)
```

```{r}
df2 %>%
  filter(trialNum==1 | trialNum==31) %>%
  spread(trialNum, data) %>%
  rename('first'=`1`, 'last'=`31`) %>%
  mutate(diff=last-first) %>%
  ggplot(aes(x=IP, y=diff)) +
    geom_bar(stat='identity', position='dodge') +
    theme_few()
```

```{r}
df2 %>%
  group_by(trialNum) %>%
  summarise(n=n(),
            avgProp=mean(data),
            ci_high=avgProp + 1.96 *sqrt(avgProp*(1-avgProp)/n),
            ci_low=avgProp - 1.96 *sqrt(avgProp*(1-avgProp)/n)) %>%
  ggplot(aes(x=trialNum, y=avgProp)) +
    geom_line() +
    geom_errorbar(aes(ymax=ci_high, ymin=ci_low)) +
    ylim(0, 1) +
    theme_few()
```

### 2019-01-07 paper/paper.Rmd
update directory structure. add paper outline.

---
title: "A pragmatic account of Zipf's Law of Abbreviation via Bayesian data analysis"
bibliography: library.bib
csl: apa6.csl
document-params: "10pt, letterpaper"

author-information: > 
    \author{{\large \bf Author 1} \\ \texttt{author1@university.edu} \\ Department of Psychology \\ Some University
    \And {\large \bf Author 2} \\ \texttt{author1@university.edu} \\ Department of Psychology \\ Some University}

abstract: 
    Zipf (1935) argued that the relation between a word's frequency and length was 
    a universal property of human languages emerging from the pressure to minimize speaker-listener effort.
    Subsequent work has provided indirect evidence for a Zipfian analysis of this relation between frequency and length, what Zipf termed
    the "Law of Abbreviation". In an effort to directly isolate the impact of speaker and listener pressures,
    Kanwal et al. (2017) conducted a study using an artificial language learning paradigm. The authors showed that 
    only when both a speaker and listener pressure were present would the Law of Abbreviation emerge. The authors
    highlight that while their data is consistent with a Zipfian analysis, it is also consistent with pragmatic langage use account.
    We consider this latter account conducting a Bayesian data analysis using the data from Kanwal et al. (2017) representing
    participants as rational pragmatic agents using the Rational Speech Act framework. Fitting participant-level parameters
    we find could fit between model posterior-predictive values on the empirical data from Kanwal et al. (2017). We believe this
    analysis provides strong evidence for a connection between Zipfian notions of efficiency driven language change and Gricean 
    notions of rational-pragmatic langauge use.

keywords:
    "Add your choice of indexing terms or keywords; kindly use a semi-colon; between each term."

output: cogsci2016::cogsci_paper
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=3, fig.height=3, fig.crop = F, fig.pos = "tb", fig.path='figs/',
                      echo=F, warning=F, cache=F, message=F, sanitize = T)
```

```{r, libraries}
library(png)
library(grid)
library(ggplot2)
library(xtable)
```

# Introduction

Zipf (1935) argued for *effort minimization* as a guiding principle of human behavior. Relying on empirical evidence from corpora, he argued that the unique dynamics of speaker- and listener-effort minimization give rise to particular distribuational forms found in natural langauge. In particular, Zipf examined the relation between a word's frequency in a corpus and a variety of properties including its rank-frequency, its denotation size, and its length. In terms of this last property, what he later termed the "Law of Abbrevation," Zipf claimed that the more frequent words are the shorter they tend to be. This relationship has been confirmed cross-linguistically, and may also be present in animal communication systems (Ferer-i-Cancho et, 2013).\par

Under a Zipfian analysis, patterns such as the "Law of Abbrevation" emerge from the competition of speaker- and listener- forces.\par

While the Zipfian analysis is compelling, other work has argued that identical distributional forms are the result of random typing (Ferrer-i-Cancho & Mascoso del Pardo, 2012).  Additionally, most evidence supporting the Zipfian approach has been indirect and did not fully explicate the causal role of competiting speaker and listener pressures.\par

Kanwal et al. (2017) implemented an artificial language learning paradigm in order to isolate these pressures directly. In line with Zipf's original analysis, they hypothesized that only in the combined speaker and listener pressure condition would the "Law of Abbreviation" emerge. That is, participants in the combined condition should converge to a lexicon in which more frequently used words were shortened and less frequent words were longer. By contrast, they hypothesized that in conditions which there was only a speaker pressure or only a listener pressure such effects would not be seen. Results largely confirmed their hypotheses.\par

While this study represents an important experimental step in confirming Zipfian notions about lexicon-level efficiency, the authors highlight an important distinction. That is, a "distinction between a language-user's mental representation of the lexicon, and the form-meaning mappings they actually produce in communication. Participants using the short form for the frequent object and the long form for the infrequent object may still retain associations of both forms with both objets in their mental lexicon-however, the nature of the communication task in this experiment may have caused them to produce only the short form for one object and the long form for the other based on purely pragmatic considerations. Given that our experiment only recorded particpnts' actual prdocutions, we cannot with certainty distinguish between these two possible explanations for the observed behavior."\par

In the current project we examine the second explanation using a Bayesian Data Analysis in which we represent study participants as rational-pragmatic speakers and listeners using the Rational Speech act framework (Frank & Goodman, 2012; Goodman & Frank, 2016). Participant-level parameters are inferred using bayesian inference. We correlate posteior-predictive values with participant-level data found by Kanwal et al. (2017) finding high degrees of agreement with our model and their data ($r^2=0.98$). While we beleive this provides compelling evidence that participants are behaving pragmatically we highlight a number of findings that our current analysis does not accomodate. Further we see these findings as possible evidence for an interpretation put forth by Kanwal et al. That is, the "pragmatics-driven asymetry in usage may or may not lead to an immediate shift in ledical representations, it may be an important first step in such a change."\par

In the following work we introduce the empirical framework of Kanwal et al. (2017) focusing on the three-conditions most relevant to our current analysis. Subsequently, we introduce the Rational Speech Act framework (RSA), our linking function in our Baysian data analysis. We provide a high-level overview of the particular modeling framework and the results of our analysis. We provide a discussion of the short-comings and implications of our approach ending with an example of how pragmatic driven asymetries may drive lexical change.\par

# Experimental evidence of Zipf's Law of Abbreviation

Previous evidence for Zipf's Law of Abbreviation was ...\par

Kanwal et al. (2017) used an artificial language learning paradigm to isolate the impact of speaker- and listener-pressures. They test four conditions. In the *combined* condition two participants were paired in web experiment....\par

The authors hypothesized that a preference to map the short form to the more frequent object, but not the infrequent object should only occur in *combined* condition. By contrast, when there was no pressure to communicate, and only the speaker cost was present, participants should always prefer the shortened form, regardless of the referrent. When there was cost different between the shortened and lengthened forms (speaker pressure) then participants shold always prefer to use the unambiguous, long-forms regardless of referrent. The authors ran an additional forth condition in which there were no speaker or listener pressures. For the purposes of the current project we did no anlyse data from this condition (REASON WHY HERE).

# A changing lexicon or pragmatic language use?

While the data is consistent with a Zipfian account of the Law of Abbreviation, Kanwal et al. (2017) highlight an important alternative explanation. They describe a distinction between the "mental representation" of the lexicon and produced forms. That is, Zipf's Law of Abbreviation is an account of lexical change, not language-use. If participants are retaining both words in their "mental lexicon," but *using the lexicon pragmatically* by using the shortened form only for the more frequent item, then the lexicon itself has not "changed" at all. Given the nature of their current task, the authors highlight that they cannot distinghish beetween the possiuble explanations.

# Modeling framework

## Rational speech act theory as a model for pragmatic participant behavior

## Plate diagram

# Results
```{r plot-post-predictives, fig.env = "figure*", fig.pos = "h", fig.width=6, fig.height=3, fig.align = "center", set.cap.width=T, num.cols.cap=2, fig.cap = "This image spans both columns. And the caption text is limited to 0.8 of the width of the document."}
img <- png::readPNG("figs/post-predictives-plot.png")
grid::grid.raster(img)
```

# Connecting pragmatic langauge use to language change

# General Formatting Instructions

For general information about authoring in markdown, see **[here](http://rmarkdown.rstudio.com/authoring_basics.html).**

For standard spoken papers and standard posters, the entire 
contribution (including figures, references, everything) can be 
no longer than six pages. For abstract posters, the entire contribution 
can be no longer than one page. For symposia, the entire contribution 
can be no longer than two pages.

The text of the paper should be formatted in two columns with an
overall width of 7 inches (17.8 cm) and length of 9.25 inches (23.5
cm), with 0.25 inches between the columns. Leave two line spaces
between the last author listed and the text of the paper. The left
margin should be 0.75 inches and the top margin should be 1 inch.
\textbf{The right and bottom margins will depend on whether you use
U.S. letter or A4 paper, so you must be sure to measure the width of
the printed text.} Use 10 point Times Roman with 12 point vertical
spacing, unless otherwise specified.

The title should be in 14 point, bold, and centered. The title should
be formatted with initial caps (the first letter of content words
capitalized and the rest lower case). Each author's name should appear
on a separate line, 11 point bold, and centered, with the author's
email address in parentheses. Under each author's name list the
author's affiliation and postal address in ordinary 10 point  type.

Indent the first line of each paragraph by 1/8~inch (except for the
first paragraph of a new section). Do not add extra vertical space
between paragraphs.

# First-Level Headings

First level headings should be in 12 point , initial caps, bold and
centered. Leave one line space above the heading and 1/4~line space
below the heading.

## Second-Level Headings

Second level headings should be 11 point , initial caps, bold, and
flush left. Leave one line space above the heading and 1/4~ line
space below the heading.

### Third-Level Headings

Third-level headings should be 10 point , initial caps, bold, and flush
left. Leave one line space above the heading, but no space after the
heading.

# Formalities, Footnotes, and Floats

Use standard APA citation format. Citations within the text should
include the author's last name and year. If the authors' names are
included in the sentence, place only the year in parentheses, as in
[-@NewellSimon1972a], but otherwise place the entire reference in
parentheses with the authors and year separated by a comma
[@NewellSimon1972a]. List multiple references alphabetically and
separate them by semicolons [@ChalnickBillman1988a; @NewellSimon1972a]. 
Use the et. al. construction only after listing all the authors to a
publication in an earlier reference and for citations with four or
more authors.

For more information on citations in RMarkdown, see **[here](http://rmarkdown.rstudio.com/authoring_bibliographies_and_citations.html#citations).**

## Footnotes

Indicate footnotes with a number\footnote{Sample of the first
footnote.} in the text. Place the footnotes in 9 point type at the
bottom of the page on which they appear. Precede the footnote with a
horizontal rule.\footnote{Sample of the second footnote.} You can also use 
markdown formatting to include footnotes using this syntax [^1].

[^1]: Sample of a markdown footnote.

## Figures

All artwork must be very dark for purposes of reproduction and should
not be hand drawn. Number figures sequentially, placing the figure
number and caption, in 10 point, after the figure with one line space
above the caption and one line space below it. If necessary, leave extra white space at
the bottom of the page to avoid splitting the figure and figure
caption. You may float figures to the top or bottom of a column, or
set wide figures across both columns.

## Two-column images

You can read local images using png package for example and plot 
it like a regular plot using grid.raster from the grid package. 
With this method you have full control of the size of your image. **Note: Image must be in .png file format for the readPNG function to work.**

You might want to display a wide figure across both columns. To do this, you change the `fig.env` chunk option to `figure*`. To align the image in the center of the page, set `fig.align` option to `center`. To format the width of your caption text, you set the `num.cols.cap` option to `2`.

```{r 2-col-image, fig.env = "figure*", fig.pos = "h", fig.width=4, fig.height=2, fig.align = "center", set.cap.width=T, num.cols.cap=2, fig.cap = "This image spans both columns. And the caption text is limited to 0.8 of the width of the document."}
img <- png::readPNG("figs/walrus.png")
grid::grid.raster(img)
```

## One-column images

Single column is the default option, but if you want set it explicitly, set `fig.env` to `figure`. Notice that the `num.cols` option for the caption width is set to `1`.

```{r image, fig.env = "figure", fig.pos = "H", fig.align='center', fig.width=2, fig.height=2, set.cap.width=T, num.cols.cap=1, fig.cap = "One column image."}
img <- png::readPNG("figs/lab_logo_stanford.png")
grid::grid.raster(img)
```

## R Plots

You can use R chunks directly to plot graphs. And you can use latex floats in the
fig.pos chunk option to have more control over the location of your plot on the page. For more information on latex placement specifiers see **[here](https://en.wikibooks.org/wiki/LaTeX/Floats,_Figures_and_Captions)**

```{r plot, fig.env="figure", fig.pos = "H", fig.align = "center", fig.width=2, fig.height=2, fig.cap = "R plot" }
x <- 0:100
y <- 2 * (x + rnorm(length(x), sd = 3) + 3)

ggplot2::ggplot(data = data.frame(x, y), 
       aes(x = x, y = y)) + 
  geom_point() + 
  geom_smooth(method = "lm")
```

## Tables

Number tables consecutively; place the table number and title (in
10 point) above the table with one line space above the caption and
one line space below it, as in Table 1. You may float
tables to the top or bottom of a column, set wide tables across both
columns.

You can use the xtable function in the xtable package.

```{r xtable, results="asis"}
n <- 100
x <- rnorm(n)
y <- 2*x + rnorm(n)
out <- lm(y ~ x)

tab1 <- xtable::xtable(summary(out)$coef, digits=c(0, 2, 2, 1, 2), 
                      caption = "This table prints across one column.")

print(tab1, type="latex", comment = F, table.placement = "H")
```

# Acknowledgements

Place acknowledgments (including funding information) in a section at
the end of the paper.

# References

```{r}
# References will be generated automatically by Pandoc and included here.
# The following code is some latex to format the bibliography. Do not remove it.
```

\setlength{\parindent}{-0.1in} 
\setlength{\leftskip}{0.125in}
\noindent

### 2019-01-07 paper/paper.Rmd
updated rd

@@
 author-information: > 
-    \author{{\large \bf Author 1} \\ \texttt{author1@university.edu} \\ Department of Psychology \\ Some University
-    \And {\large \bf Author 2} \\ \texttt{author1@university.edu} \\ Department of Psychology \\ Some University}
+    \author{{\large \bf Benjamin N. Peloquin} \\ \texttt{bpeloqui@stanford.edu} \\ Department of Psychology \\ Stanford University
+    \And
+    {\large \bf Noah D. Goodman} \\ \texttt{noah@university.edu} \\ Department of Computer Science \\ Stanford University
+    \And {\large \bf Michael C. Frank} \\ \texttt{mike@university.edu} \\ Department of Psychology \\ Stanford University}
 
@@
 
-Zipf (1935) argued for *effort minimization* as a guiding principle of human behavior.
-Relying on empirical evidence from corpora, he argued that the unique dynamics of speaker- and listener-effort minimization give rise to particular distribuational forms found in natural langauge.
-In particular, Zipf examined the relation between a word's frequency in a corpus and a variety of properties including its rank-frequency, its denotation size, and its length.
-In terms of this last property, what he later termed the "Law of Abbrevation," Zipf claimed that the more frequent words are the shorter they tend to be.
-This relationship has been confirmed cross-linguistically, and may also be present in animal communication systems (Ferer-i-Cancho et, 2013).\par
+Zipf (1935) presented a view of human behavior that focused on  *effort minimization* as a guiding principle (Principle of Least Effort).
+Relying on empirical evidence from corpora, he argued that the unique dynamics of speaker- and listener-effort minimization give rise to particular distributional forms found in natural language.
+Zipf examined the relation between a word's frequency in a corpus and a variety of properties including its rank-frequency, its denotation size, and its length.
+In terms of this last property, what he later termed the "Law of Abbreviation," Zipf claimed that the more frequent words are the shorter they tend to be.
+This relationship has been confirmed cross-linguistically, and may also be present in animal communication systems (Ferrer-i-Cancho et, 2013).\par
 
-Under a Zipfian analysis, patterns such as the "Law of Abbrevation" emerge from the competition of speaker- and listener- forces.\par
+Zipf argued that this relationship between frequency and length was an emergent property of competing speaker-listener pressures.
+Under this framing, communication can be considered a cooperative, joint activity between interlocutors.
+Importantly, however, this activity contains a fundamental asymmetry in terms of what is effortful for a speaker compare to what is effortful for a listener.
+An optimal language, from a speaker perspective, what Zipf termed “Speaker’s economy,” would consist of a single, low-cost word which was fully ambiguous - referring to any possible referent in the world.
+In using such a language a listener would need to disambiguate every utterance by a speaker.
+By contrast, if a language were optimised solely in terms of listener effort, what Zipf termed “Auditor’s economy,” the language should bijectively map all forms to meanings so there was no need for a listener to disambiguate.\par
 
-While the Zipfian analysis is compelling, other work has argued that identical distributional forms are the result of random typing (Ferrer-i-Cancho & Mascoso del Pardo, 2012).
-Additionally, most evidence supporting the Zipfian approach has been indirect and did not fully explicate the causal role of competiting speaker and listener pressures.\par
+Zipf argued that these competing forces -- Speaker’s and Auditor’s economy -- give rise to the particular distributional forms found in all languages, including the particular relationship between word frequency and length.
+The issue is far from settled, however -- subsequent work has argued that such properties could also be explained by a process of random typing (Ferrer-i-Cancho & Mascoso del Pardo, 2012).
+Moreover, most evidence supporting the Zipfian approach has been indirect and did not fully determine the causal role of competing speaker and listener pressures.\par
 
@@
 That is, a "distinction between a language-user's mental representation of the lexicon, and the form-meaning mappings they actually produce in communication.
-Participants using the short form for the frequent object and the long form for the infrequent object may still retain associations of both forms with both objets in their mental lexicon-however, the nature of the communication task in this experiment may have caused them to produce only the short form for one object and the long form for the other based on purely pragmatic considerations.
-Given that our experiment only recorded particpnts' actual prdocutions, we cannot with certainty distinguish between these two possible explanations for the observed behavior."\par
+Participants using the short form for the frequent object and the long form for the infrequent object may still retain associations of both forms with both objects in their mental lexicon-however, the nature of the communication task in this experiment may have caused them to produce only the short form for one object and the long form for the other based on purely pragmatic considerations.
+Given that our experiment only recorded participants’ actual productions, we cannot with certainty distinguish between these two possible explanations for the observed behavior."\par
 
@@
 Participant-level parameters are inferred using bayesian inference.
-We correlate posteior-predictive values with participant-level data found by Kanwal et al.
+We correlate posterior-predictive values with participant-level data found by Kanwal et al.
 (2017) finding high degrees of agreement with our model and their data ($r^2=0.98$).
-While we beleive this provides compelling evidence that participants are behaving pragmatically we highlight a number of findings that our current analysis does not accomodate.
+While we believe this provides compelling evidence that participants are behaving pragmatically we highlight a number of findings that our current analysis does not accomodate.
 Further we see these findings as possible evidence for an interpretation put forth by Kanwal et al.
-That is, the "pragmatics-driven asymetry in usage may or may not lead to an immediate shift in ledical representations, it may be an important first step in such a change."\par
+That is, the "pragmatics-driven asymmetry in usage may or may not lead to an immediate shift in lexical representations, it may be an important first step in such a change."\par
 
@@
 (2017) focusing on the three-conditions most relevant to our current analysis.
-Subsequently, we introduce the Rational Speech Act framework (RSA), our linking function in our Baysian data analysis.
+Subsequently, we introduce the Rational Speech Act framework (RSA), our linking function in our Bayesian data analysis.
 We provide a high-level overview of the particular modeling framework and the results of our analysis.
-We provide a discussion of the short-comings and implications of our approach ending with an example of how pragmatic driven asymetries may drive lexical change.\par
+We provide a discussion of the short-comings and implications of our approach ending with an example of how pragmatic driven asymmetries may drive lexical change.\par
 
@@
 The authors hypothesized that a preference to map the short form to the more frequent object, but not the infrequent object should only occur in *combined* condition.
-By contrast, when there was no pressure to communicate, and only the speaker cost was present, participants should always prefer the shortened form, regardless of the referrent.
-When there was cost different between the shortened and lengthened forms (speaker pressure) then participants shold always prefer to use the unambiguous, long-forms regardless of referrent.
-The authors ran an additional forth condition in which there were no speaker or listener pressures.
-For the purposes of the current project we did no anlyse data from this condition (REASON WHY HERE).
+By contrast, when there was no pressure to communicate, and only the speaker cost was present, participants should always prefer the shortened form, regardless of the referent.
+When there was cost different between the shortened and lengthened forms (speaker pressure) then participants should always prefer to use the unambiguous, long-forms regardless of referent.
+The authors ran an additional fourth condition in which there were no speaker or listener pressures.
+For the purposes of the current project we did no analyse data from this condition (REASON WHY HERE).
 
@@
 If participants are retaining both words in their "mental lexicon," but *using the lexicon pragmatically* by using the shortened form only for the more frequent item, then the lexicon itself has not "changed" at all.
-Given the nature of their current task, the authors highlight that they cannot distinghish beetween the possiuble explanations.
+Given the nature of their current task, the authors highlight that they cannot distinguish between the possible explanations.
 
@@
 
-## Plate diagram
+```{r plate-diagram, fig.env = "figure", fig.pos = "H", fig.align='center', fig.width=2, fig.height=2, set.cap.width=T, num.cols.cap=1, fig.cap = "One column image."}
+img <- png::readPNG("figs/plate-temp.png")
+grid::grid.raster(img)
+```
 
@@
 
-# Connecting pragmatic langauge use to language change
+# Connecting pragmatic language use to language change
 
-# General Formatting Instructions
+# Conclusion
 
-For general information about authoring in markdown, see **[here](http://rmarkdown.rstudio.com/authoring_basics.html).**
-
-For standard spoken papers and standard posters, the entire 
-contribution (including figures, references, everything) can be 
-no longer than six pages. For abstract posters, the entire contribution 
-can be no longer than one page. For symposia, the entire contribution 
-can be no longer than two pages.
-
-The text of the paper should be formatted in two columns with an
-overall width of 7 inches (17.8 cm) and length of 9.25 inches (23.5
-cm), with 0.25 inches between the columns. Leave two line spaces
-between the last author listed and the text of the paper. The left
-margin should be 0.75 inches and the top margin should be 1 inch.
-\textbf{The right and bottom margins will depend on whether you use
-U.S. letter or A4 paper, so you must be sure to measure the width of
-the printed text.} Use 10 point Times Roman with 12 point vertical
-spacing, unless otherwise specified.
-
-The title should be in 14 point, bold, and centered. The title should
-be formatted with initial caps (the first letter of content words
-capitalized and the rest lower case). Each author's name should appear
-on a separate line, 11 point bold, and centered, with the author's
-email address in parentheses. Under each author's name list the
-author's affiliation and postal address in ordinary 10 point  type.
-
-Indent the first line of each paragraph by 1/8~inch (except for the
-first paragraph of a new section). Do not add extra vertical space
-between paragraphs.
-
-# First-Level Headings
-
-First level headings should be in 12 point , initial caps, bold and
-centered. Leave one line space above the heading and 1/4~line space
-below the heading.
-
-## Second-Level Headings
-
-Second level headings should be 11 point , initial caps, bold, and
-flush left. Leave one line space above the heading and 1/4~ line
-space below the heading.
-
-### Third-Level Headings
-
-Third-level headings should be 10 point , initial caps, bold, and flush
-left. Leave one line space above the heading, but no space after the
-heading.
-
-# Formalities, Footnotes, and Floats
-
-Use standard APA citation format. Citations within the text should
-include the author's last name and year. If the authors' names are
-included in the sentence, place only the year in parentheses, as in
-[-@NewellSimon1972a], but otherwise place the entire reference in
-parentheses with the authors and year separated by a comma
-[@NewellSimon1972a]. List multiple references alphabetically and
-separate them by semicolons [@ChalnickBillman1988a; @NewellSimon1972a]. 
-Use the et. al. construction only after listing all the authors to a
-publication in an earlier reference and for citations with four or
-more authors.
-
-For more information on citations in RMarkdown, see **[here](http://rmarkdown.rstudio.com/authoring_bibliographies_and_citations.html#citations).**
-
-## Footnotes
-
-Indicate footnotes with a number\footnote{Sample of the first
-footnote.} in the text. Place the footnotes in 9 point type at the
-bottom of the page on which they appear. Precede the footnote with a
-horizontal rule.\footnote{Sample of the second footnote.} You can also use 
-markdown formatting to include footnotes using this syntax [^1].
-
-[^1]: Sample of a markdown footnote.
-
-## Figures
-
-All artwork must be very dark for purposes of reproduction and should
-not be hand drawn. Number figures sequentially, placing the figure
-number and caption, in 10 point, after the figure with one line space
-above the caption and one line space below it. If necessary, leave extra white space at
-the bottom of the page to avoid splitting the figure and figure
-caption. You may float figures to the top or bottom of a column, or
-set wide figures across both columns.
-
-## Two-column images
-
-You can read local images using png package for example and plot 
-it like a regular plot using grid.raster from the grid package. 
-With this method you have full control of the size of your image. **Note: Image must be in .png file format for the readPNG function to work.**
-
-You might want to display a wide figure across both columns.
-To do this, you change the `fig.env` chunk option to `figure*`.
-To align the image in the center of the page, set `fig.align` option to `center`.
-To format the width of your caption text, you set the `num.cols.cap` option to `2`.
-
-```{r 2-col-image, fig.env = "figure*", fig.pos = "h", fig.width=4, fig.height=2, fig.align = "center", set.cap.width=T, num.cols.cap=2, fig.cap = "This image spans both columns.
-And the caption text is limited to 0.8 of the width of the document."}
-img <- png::readPNG("figs/walrus.png")
-grid::grid.raster(img)
-```
-
-## One-column images
-
-Single column is the default option, but if you want set it explicitly, set `fig.env` to `figure`. Notice that the `num.cols` option for the caption width is set to `1`.
-
-```{r image, fig.env = "figure", fig.pos = "H", fig.align='center', fig.width=2, fig.height=2, set.cap.width=T, num.cols.cap=1, fig.cap = "One column image."}
-img <- png::readPNG("figs/lab_logo_stanford.png")
-grid::grid.raster(img)
-```
-
-## R Plots
-
-You can use R chunks directly to plot graphs. And you can use latex floats in the
-fig.pos chunk option to have more control over the location of your plot on the page.
-For more information on latex placement specifiers see **[here](https://en.wikibooks.org/wiki/LaTeX/Floats,_Figures_and_Captions)**
-
-```{r plot, fig.env="figure", fig.pos = "H", fig.align = "center", fig.width=2, fig.height=2, fig.cap = "R plot" }
-x <- 0:100
-y <- 2 * (x + rnorm(length(x), sd = 3) + 3)
-
-ggplot2::ggplot(data = data.frame(x, y), 
-       aes(x = x, y = y)) + 
-  geom_point() + 
-  geom_smooth(method = "lm")
-```
-
-## Tables
-
-Number tables consecutively; place the table number and title (in
-10 point) above the table with one line space above the caption and
-one line space below it, as in Table 1. You may float
-tables to the top or bottom of a column, set wide tables across both
-columns.
-
-You can use the xtable function in the xtable package.
-
-```{r xtable, results="asis"}
-n <- 100
-x <- rnorm(n)
-y <- 2*x + rnorm(n)
-out <- lm(y ~ x)
-
-tab1 <- xtable::xtable(summary(out)$coef, digits=c(0, 2, 2, 1, 2), 
-                      caption = "This table prints across one column.")
-
-print(tab1, type="latex", comment = F, table.placement = "H")
-```
+Languages around the world display a consistent relationship between word length and frequency -- more frequent words tend to be shorter (CITATIONS).
+Zipf (1935) called this particular finding the “Law of Abbreviation” and characterized it in terms of competing speaker listener pressures.
+While other work has provided indirect evidence consistent with this Zipfian analysis (CITATIONS), Kanwal et al.
+(2017) attempted to derive this effect while isolating speaker and listener pressures experimentally.
+In an artificial language learning paradigm the authors showed that subjects were significantly more likely to produce lexicons consistent the “Law of Abbreviation,” but only when both speaker and listener pressures are present.
+While this evidence is compelling the authors highlight an alternative hypothesis that may have generated the same effect, but could not be distinguished in the current study -- pragmatic language use instead of lexical change.
+To test this secondary hypothesis we conducted a Bayesian data analysis, modeling participants as rational pragmatic agents using the Rational Speech-act (RSA) framework.
+Inferring subject-level parameters we are able to fit Kanwal et al.
+(2017) experimental data with high-levels of accuracy ($r^2=0.986).
+While we believe this provides evidence for the pragmatic-language use interpretation we also highlight the compatibility of this interpretation with an additional insight of Kanwal et al.
+(2017) -- pragmatic language use may serve as an important bootstrap for future language change.
 

### 2019-01-07 bda/analysis/fit_model.Rmd
updated plotting -- model fitting vs baseline + inferred subj param distributions

@@
 library(ggthemes)
+library(gridExtra)
 library(rwebppl)
@@
 ```{r}
-f_bda_model <- "../models/bda.wppl"
-model_bda <- getModelFile(f_bda_model)
+# f_bda_model <- "../models/bda.wppl"
+# model_bda <- getModelFile(f_bda_model)
 ```
@@
 ```{r}
-problemIPs <- c('142.196.192.44', 
-                '173.242.228.167',
-                '104.57.161.213',
-                '104.11.62.66',
-                '184.155.134.46',
-                '67.168.183.190',
-                '108.210.65.252')
+# problemIPs <- c('142.196.192.44', 
+#                 '173.242.228.167',
+#                 '104.57.161.213',
+#                 '104.11.62.66',
+#                 '184.155.134.46',
+#                 '67.168.183.190',
+#                 '108.210.65.252')
 ```
@@
 ```{r}
+plot_all_post_preds <- df_predictives2 %>%
+  # mutate(actualFrequentProps=frequent_cnt / frequent_n,
+  #        actualInrequentProps=infrequent_cnt / infrequent_n,
+  #        estFrequentProps=posteriorPr5edictiveFrequent / frequent_n,
+  #        estInfrequentProps=posteriorPredictiveInfrequent / infrequent_n) %>%
+  # select(IP, condition, inferredCondition, actualFrequentProps, estFrequentProps, actualInrequentProps, estInfrequentProps) %>%
+  # gather(type, value, c(posteriorPredictiveFrequent, posteriorPredictiveInfrequent))
+  ggplot(aes(x=actualCnt, y=estCnt, col=condition), alpha=0.4) +
+    geom_smooth(aes(group=0), method='lm') +
+    geom_point(aes(fill=condition), alpha=0.6, size=3, colour="white", pch=21) +
+    # geom_point(aes(x=actualInrequentProps, y=estInfrequentProps, col=condition), alpha=0.4) +
+    ylab('Posterior predictive \n counts of short form') +
+    xlab('Human data \n counts of short form') +
+    ggtitle('r^2=0.987') +
+    theme_few() +
+    theme(plot.title = element_text(hjust = 0.5))
+
+ggsave('../../paper/figs/all-post-predictives-plot.png', plot_all_post_preds, width=6, height=3)
+```
+
+```{r}
 df_predictives2 %>%
@@
 
-ggsave('../../paper/figs/post-predictives-plot.png', width=6, height=3)
+ggsave('../../paper/figs/post-predictives-facet-plot.png', width=6, height=3)
 ```
@@
 ```{r}
-df_predictives2 %>%
+bda_vs_baseline_plot <- df_predictives2 %>%
   mutate(baseline=ifelse(condition=='combined' & type=='frequent', 24,
@@
                                               ifelse(condition=='time' & type=='frequent', 48, 16)))))) %>%
-  ggplot(aes(x=actualCnt, y=baseline, col=condition), alpha=0.4) +
+  gather('model', 'pred', c(baseline, estCnt)) %>%
+  mutate(model=ifelse(model=='estCnt', 'BDA', 'Basline')) %>%
+  ggplot(aes(x=actualCnt, y=pred, col=condition), alpha=0.4) +
     geom_point(aes(fill=condition), alpha=0.6, size=3, colour="white", pch=21) +
+    geom_abline(intercept=0, slope=1, col='grey', alpha=0.4) +
     # geom_point(aes(x=actualInrequentProps, y=estInfrequentProps, col=condition), alpha=0.4) +
-    facet_wrap(~condition, scales="free") +
+    facet_wrap(~model, scales="free") +
     ylab('Baseline model \n counts of short form') +
@@
 
+ggsave('../../paper/figs/bda-vs-baseline-plot.png', bda_vs_baseline_plot, width=6, height=3)
+```
+
+```{r}
 df_predictives2 %>%
@@
   summarise(r2 = cor(actualCnt, baseline))
+
+```
+
+## Subject-level inferred params?
+
+```{r}
+plot_alpha <- df_predictives %>%
+  ggplot(aes(alpha)) +
+    geom_density() +
+    theme_few()
+
+plot_depth <- df_predictives %>%
+  ggplot(aes(depth)) +
+    geom_histogram(bins=3) +
+    theme_few()
+
+plot_attention <- df_predictives %>%
+  ggplot(aes(inferredAttention)) +
+    geom_density() +
+    theme_few()
+
+plot_condition <- df_predictives %>%
+  mutate(inferredCondition=as.character(inferredCondition)) %>%
+  mutate(inferredCondition=ifelse(inferredCondition=='conventionalized', 'other', inferredCondition)) %>%
+  group_by(inferredCondition) %>%
+  summarise(count=n()) %>%
+  ggplot(aes(x=inferredCondition, y=count)) +
+    geom_bar(stat='identity', position='dodge') +
+    theme_few() +
+    theme(axis.text.x = element_text(angle = 45, hjust = 1))
+
+plot_all_params <- grid.arrange(plot_alpha, plot_depth, plot_attention, plot_condition, nrow=2)
+ggsave('../../paper/figs/inferred-params-plot.png', plot=plot_all_params, width=6, height=4)
+# plot_condition <- df_predictives %>%
+#   group_by(inferredCondition) %>%
+#   summarise(n=n())
 ```
@@
 mean(df_predictives$inferredCondition == df_predictives$actualCondition)
+```
+
+We use conventionatlized only for a few. These are very strange in that they *almost* always avoid using the short form for the frequent term, but use it for the non-frequent 
+```{r}
+df_predictives %>%
+  filter(inferredCondition=='conventionalized')
 ```

### 2019-01-07 paper/paper.Rmd
updates to paper.

@@
     Kanwal et al. (2017) conducted a study using an artificial language learning paradigm. The authors showed that 
-    only when both a speaker and listener pressure were present would the Law of Abbreviation emerge. The authors
+    only when both a speaker and listener pressure were present did languages consistent with the Law of Abbreviation emerge. The authors
     highlight that while their data is consistent with a Zipfian analysis, it is also consistent with pragmatic langage use account.
     We consider this latter account conducting a Bayesian data analysis using the data from Kanwal et al. (2017) representing
-    participants as rational pragmatic agents using the Rational Speech Act framework. Fitting participant-level parameters
-    we find could fit between model posterior-predictive values on the empirical data from Kanwal et al. (2017). We believe this
-    analysis provides strong evidence for a connection between Zipfian notions of efficiency driven language change and Gricean 
+    subjects as rational, pragmatic agents using the Rational Speech Act framework. Fitting subject-level parameters
+    we find good fit between model posterior-predictive values and data from Kanwal et al. (2017). We argue this
+    analysis provides evidence for a connection between Zipfian notions of efficiency driven language change and Gricean 
     notions of rational-pragmatic langauge use.
@@
 
-Zipf (1935) presented a view of human behavior that focused on  *effort minimization* as a guiding principle (Principle of Least Effort).
-Relying on empirical evidence from corpora, he argued that the unique dynamics of speaker- and listener-effort minimization give rise to particular distributional forms found in natural language.
-Zipf examined the relation between a word's frequency in a corpus and a variety of properties including its rank-frequency, its denotation size, and its length.
-In terms of this last property, what he later termed the "Law of Abbreviation," Zipf claimed that the more frequent words are the shorter they tend to be.
-This relationship has been confirmed cross-linguistically, and may also be present in animal communication systems (Ferrer-i-Cancho et, 2013).\par
+Zipf (1935) presented a view of human behavior that focused on *effort minimization* as a guiding principle.
+Relying on empirical evidence from corpora, he argued that the unique dynamics of speaker- and listener-effort minimization give rise to the distributional forms found  in natural language.
+Among these, Zipf examined the relation between a word's frequency in a corpus and a variety of properties including its rank-frequency, its denotation size, and its length.
+In terms of this last property, what he later termed the "Law of Abbreviation," Zipf claimed that the more frequent words are, the shorter they tend to be.
+Evidence for this relationship can be found across languages, may also be present in animal communication systems  (Ferrer-i-Cancho et, 2013), and is a property of artificial (programming) languages as well.\par
 
 Zipf argued that this relationship between frequency and length was an emergent property of competing speaker-listener pressures.
-Under this framing, communication can be considered a cooperative, joint activity between interlocutors.
-Importantly, however, this activity contains a fundamental asymmetry in terms of what is effortful for a speaker compare to what is effortful for a listener.
-An optimal language, from a speaker perspective, what Zipf termed “Speaker’s economy,” would consist of a single, low-cost word which was fully ambiguous - referring to any possible referent in the world.
-In using such a language a listener would need to disambiguate every utterance by a speaker.
-By contrast, if a language were optimised solely in terms of listener effort, what Zipf termed “Auditor’s economy,” the language should bijectively map all forms to meanings so there was no need for a listener to disambiguate.\par
+Under this framing, communication is viewed as a cooperative, joint activity between interlocutors.
+Importantly, there remains a fundamental asymmetry in any communicative act -- what is effortful for a speaker (production) is different than what is effortful for a listener (understanding).
+Zipf suggested that an optimal language from a speaker perspective would consist of a single, low-cost word which was fully ambiguous - referring to any possible referent in the world.
+In using such a language a listener would need to disambiguate every utterance by the speaker.
+By contrast, if a language were optimised solely in terms of listener effort, the language should bijectively map all forms to meanings so there was no need for a listener to disambiguate.\par
 
-Zipf argued that these competing forces -- Speaker’s and Auditor’s economy -- give rise to the particular distributional forms found in all languages, including the particular relationship between word frequency and length.
-The issue is far from settled, however -- subsequent work has argued that such properties could also be explained by a process of random typing (Ferrer-i-Cancho & Mascoso del Pardo, 2012).
+Zipf argued that these competing forces, what he called “Speaker’s economy” and “Auditor’s economy”, give rise to the particular distributional forms found in all languages, including the relationship between word frequency and length.
+While this framework is compelling and can largely be seen as foundational to a family of theories examining efficiency in language-structure (CITATIONS) and  use (CITATIONS), the issue is far from settled.
+Subsequent work has argued that properties such as the relation between word frequency and length could also be explained by a process of random typing (Ferrer-i-Cancho & Mascoso del Pardo, 2012).
 Moreover, most evidence supporting the Zipfian approach has been indirect and did not fully determine the causal role of competing speaker and listener pressures.\par
@@
 Kanwal et al.
-(2017) implemented an artificial language learning paradigm in order to isolate these pressures directly.
-In line with Zipf's original analysis, they hypothesized that only in the combined speaker and listener pressure condition would the "Law of Abbreviation" emerge.
-That is, participants in the combined condition should converge to a lexicon in which more frequently used words were shortened and less frequent words were longer.
-By contrast, they hypothesized that in conditions which there was only a speaker pressure or only a listener pressure such effects would not be seen.
+(2017) implemented an artificial language learning paradigm to isolate speaker-listener pressures directly.
+In line with Zipf's original analysis, they hypothesized that only when *both* speaker and listener pressures are present in a communication setting would the "Law of Abbreviation" emerge.
+The authors adopted a reference game setting in which participants learned a small language consisting of three words, which could be used to refer to two items.
+They hypothesized that when given a choice between words subjects should prefer to use the less costly, but ambiguous term to refer to a more frequently occurring object.
+Importantly, however, this choice should only appear when there was both a speaker- and listener-pressures present.
 Results largely confirmed their hypotheses.\par
 
-While this study represents an important experimental step in confirming Zipfian notions about lexicon-level efficiency, the authors highlight an important distinction.
-That is, a "distinction between a language-user's mental representation of the lexicon, and the form-meaning mappings they actually produce in communication.
-Participants using the short form for the frequent object and the long form for the infrequent object may still retain associations of both forms with both objects in their mental lexicon-however, the nature of the communication task in this experiment may have caused them to produce only the short form for one object and the long form for the other based on purely pragmatic considerations.
-Given that our experiment only recorded participants’ actual productions, we cannot with certainty distinguish between these two possible explanations for the observed behavior."\par
+While Kanwal et al.
+(2017) represents an important experimental step in confirming Zipfian notions about lexicon-level efficiency, the authors highlight an important and subtle dimension to these results.
+Zipf’s “Law of Abbreviation” is a lexicon-level theory.
+That is, it attempts to explain why particular word forms are mapped to particular meanings.
+However, their results are, in theory, consistent with a pragmatic language-use account.\par
 
-In the current project we examine the second explanation using a Bayesian Data Analysis in which we represent study participants as rational-pragmatic speakers and listeners using the Rational Speech act framework (Frank & Goodman, 2012; Goodman & Frank, 2016).
-Participant-level parameters are inferred using bayesian inference.
-We correlate posterior-predictive values with participant-level data found by Kanwal et al.
-(2017) finding high degrees of agreement with our model and their data ($r^2=0.98$).
-While we believe this provides compelling evidence that participants are behaving pragmatically we highlight a number of findings that our current analysis does not accomodate.
-Further we see these findings as possible evidence for an interpretation put forth by Kanwal et al.
-That is, the "pragmatics-driven asymmetry in usage may or may not lead to an immediate shift in lexical representations, it may be an important first step in such a change."\par
+*"There is a distinction between a language-user's mental representation of the lexicon, and the form-meaning mappings they actually produce in communication...the nature of the communication task in this experiment may have caused them to produce only the short form for one object and the long form for the other based on purely pragmatic considerations."*\par
 
-In the following work we introduce the empirical framework of Kanwal et al.
-(2017) focusing on the three-conditions most relevant to our current analysis.
-Subsequently, we introduce the Rational Speech Act framework (RSA), our linking function in our Bayesian data analysis.
-We provide a high-level overview of the particular modeling framework and the results of our analysis.
-We provide a discussion of the short-comings and implications of our approach ending with an example of how pragmatic driven asymmetries may drive lexical change.\par
+Put differently, the underlying lexicon learned by subjects may have remained unchanged during the experiment -- a finding which, on its face, is at odds with the claim that this experiment demonstrates lexicon-level change.
+Rather, subjects may have *used* the lexicon pragmatically, preferring to map less costly forms to more frequent objects, while leaving the underlying lexicon unchanged.
+The authors’ correctly point out that because they only recorded participants’ actual language production they cannot distinguishing between the lexicon change account and pragmatic accounts.
+
+The current project addresses the question of whether the patterns of production behavior observed in Kanwal et al.
+(2017) can be explained by a pragmatic, language-use  account.
+To do so, we conduct a Bayesian data analysis, modeling subjects as rational pragmatic agents using the Rational Speech-Act framework (RSA) (Frank & Goodman, 2017; Goodman & Frank, 2016).
+To preview our results we find good fit from model posterior-predictive values to Kanwal et al.
+(2017) participant data ($R^2=0.986$), over and above what would would be expected from a baseline model ($r^2=0.68$).
+We believe this provides compelling evidence in support of the alternative hypothesis outlined by Kanwal et al.
+(2017) -- that subjects may have been behaving pragmatically, rather than demonstrating real change at the level of language-structure.
+\par
+
+The paper proceeds as follows.
+We describe the experimental set-up of Kanwal et al.
+(2017), their hypotheses and the alternative interpretation of their results.
+We then introduce our general approach to the Bayesian data analysis as well as the Rational Speech-act framework we will use to model pragmatic language use.
+We evaluate our results summarising the findings and highlighting aspects that our modeling does not capture.
+Following this work we return the question posed by Kanwal et al.
+(2017) of how pragmatic language use might bootstrap language change and offer an argument for how this type of process could in theory unfold.
+We end with a summary of our findings and next steps.\par
 
@@
 
-Previous evidence for Zipf's Law of Abbreviation was ...\par
+Kanwal et al.
+(2017) used an artificial language learning paradigm to isolate the impact of speaker- and listener-pressures on language structure.
+During training, participants first learn a miniature language.
+During test, participant’s language use is recorded, typically over a series of communication events or across multiple generations of participants.
+This paradigm has been used to study changes in language structure (CITATIONS).\par
 
 Kanwal et al.
-(2017) used an artificial language learning paradigm to isolate the impact of speaker- and listener-pressures.
-They test four conditions.
-In the *combined* condition two participants were paired in web experiment....\par
+(2017) adopt this basic framework -- participants learned a miniature artificial language and were asked to communicate with an interlocutor in a reference game setting.
+Following ideas first indicated by Zipf (1935), the authors hypothesized the presence or absence of pressures to communicate quickly (speaker pressure) and accurately (listener pressure)  should impact the resulting languages.
+The study proceeded as follows.
+Participants were assigned to one of four possible conditions.
+A *combined* condition in which both speaker- and listener-pressures were present.
+An *accuracy* condition, in which only a listener pressure was present.
+A *time* condition in which only a speaker pressure was present.
+They also tested a *neither* condition in which neither speaker nor listener pressures were present.
+For the purpose of our current project, we do not model the *neither* condition as it's unclear what optimal behavior should like.\par
+
+For a comprehensive description of the experimental set-up we refer the reader to Kanwal et al.
+(2017), focusing on essential details in this section.
+During training participants learned labels for two novel objects.
+Importantly, one of the objects occurred more frequently than the other.
+During training participants observed the objects at their relative frequencies as well as labels.
+Labels appeared either as a long-form (e.g.
+“zopudon” or “zopekil”) or short form (“zop”).
+Importantly, in this set-up “zop” was ambiguous between the two object referents.
+During test, pairs of participants were asked to participate in a communication game in which a “director” was shown an object” and asked to transmit its name to the “matcher.” The “director” was always given a choice to use the long-form label or the ambiguous short-form.
+Crucially, to transmit the label to the matcher the director had to click on it, holding the mouse while each character appeared sequentially.
+In this setting, transmitting a longer-form required more time -- an elegant operationalization of speaker production cost.
+Listener pressure was operationalized via a pressure for the matcher to correctly identify the referent.\par
 
 The authors hypothesized that a preference to map the short form to the more frequent object, but not the infrequent object should only occur in *combined* condition.
-By contrast, when there was no pressure to communicate, and only the speaker cost was present, participants should always prefer the shortened form, regardless of the referent.
-When there was cost different between the shortened and lengthened forms (speaker pressure) then participants should always prefer to use the unambiguous, long-forms regardless of referent.
-The authors ran an additional fourth condition in which there were no speaker or listener pressures.
-For the purposes of the current project we did no analyse data from this condition (REASON WHY HERE).
+By contrast, when there was no pressure to communicate (no listener was present in the *time* condition), and only the speaker cost was present, participants should always prefer the shortened form, regardless of the referent.
+When there was no cost difference between the shortened and lengthened forms, but a listener was present, then participants should always prefer to use the unambiguous, long-forms regardless of referentin the *accuracy* condition.
+Results were largely consistent with these hypotheses.
 
@@
 If participants are retaining both words in their "mental lexicon," but *using the lexicon pragmatically* by using the shortened form only for the more frequent item, then the lexicon itself has not "changed" at all.
-Given the nature of their current task, the authors highlight that they cannot distinguish between the possible explanations.
+This analysis falls largely in line with Gricean notion of natural language pragmatics.\par
+
+Grice (1957) argued that communication could largely be analysed as an instance of rational, cooperative behavior.
+Under this framework speakers choose utterances to convey meanings.
+Listeners attempt to infer the speaker's communicative goal.
+At the heart of Grice’s theory was a set of conversational maxims known to both speakers and listeners (be truthful, relevant, informative and perspicuous).
+Inferences about a speaker’s intended meaning could be derived from these maxims by a listener.
+In particular, it is assumed that the listener can reason about the speaker and information that is mutually known to both.
+The matcher can use such information, such as the fact that using the short-form is less costly than using a long-form, and that the frequent object occurs three times more often than the infrequent object.
+The matcher and director might reason as follows: *If the director is trying to transmit information as quickly as possible they should always use the quicker, shortened form.
+But if they want me to pick out the correct referent they should always transmit the slower, longer form.
+An optimal solution should use the short form in in as many trials as possible.
+Therefore we should map the shorter, ambiguous form to the more frequent object.*\par
+
+Notably, by reasoning about one-another, about the goals of the game (be quick, be accurate), and the details of context (differential word costs and object frequencies) a pair can converge on optimal *use* of the provided language, but *without changing the underlying structure*.
+
+## Rational speech act theory as a model for pragmatic participant behavior
+
+To model this type of rational, pragmatic interaction we adopt the Rational Speech Act Framework (Frank & Goodman, 2012; Goodman & Frank, 2016).
+RSA is a recursive Bayesian model of pragmatic language production and interpretation, which can largely be seen as a mathematical formalization of essential Gricean principles.
+RSA has proven to be a productive framework for modeling a range of pragmatic phenomena from hyperbole, metaphor, scalar implicature and others (CITATIONS).
+Under RSA a speaker agent describes a conditional distribution mapping meanings $u\in U$ to utterances $m\in M$, written as $S(u|m)$.
+A speaker agent describes a conditional distribution mapping from utterances to meaning, written as $L(m|u)$.
+Importantly, each of these functions is described in terms of the other.
+That is,
+
+\begin{equation}
+S_i(u|m) \propto e^{U(u;m)}
+\end{equation}
+Where
+\begin{equation}
+U(u;m) = \alpha(log(L_{i-1}(m|u)) - cost(u))
+\end{equation}
+And 
+\begin{equation}
+L_{i-1}(m|u) \propto S_{i-1}(u|m)\times p(m)
+\end{equation}
+Defining nested speaker and listener agents could, in principle lead to infinite recursion.
+We take a *literal listener*, denoted $L_0(m|u) as a basecase.
+The literal listener does not reason about a speaker model, rather this agent considers the literal semantics of the utterance.
+\begin{equation}
+L_0(m|u) \propto \delta_u(m)\times p(m)
+\end{equation}
+Where
+\begin{equation}
+\delta_u(m)=
+\begin{cases}
+1, & \text{if } m \in [[u]]\\
+0, & o.w.
+\end{cases}
+\end{equation}
 
@@
 
-## Rational speech act theory as a model for pragmatic participant behavior
+Top assess the degree to which the observed production data in Kanwal et al.
+(2017) can be described by a pragmatic language-use account we conduct a Bayesian data analysis, using RSA as our linking function.
+Our analysis proceeds in two parts.
+In the parameter inference phase we infer subject-level parameters.
+In the model-checking and posterior prediction phase we assess how well our model, with the inferred parameters, describes the experimental data.
 
-```{r plate-diagram, fig.env = "figure", fig.pos = "H", fig.align='center', fig.width=2, fig.height=2, set.cap.width=T, num.cols.cap=1, fig.cap = "One column image."}
+### Parameter inference
+
+For each participant we infer a set of parameters -- $\alpha$, $\beta$ are RSA model parameters corresponding the the subjects rationality and recursive depth, respectively.
+The parameters $\gamma$ and $\lambda$ correspond to the particular condition (made up of utterance costs and object referent priors) as wells as a focus parameter describing the probability that the subject is not paying attention and choosing to produce utterance randomly.
+Table 1 includes the parameters and their descriptions.\par
+
+```{r plate-notation, fig.env = "figure", fig.pos = "H", fig.align='center', fig.width=2, fig.height=2, set.cap.width=T, num.cols.cap=1, fig.cap = "One column image."}
 img <- png::readPNG("figs/plate-temp.png")
@@
 
+For each subject we use $n=200$ samples using Metropolis-Hasting to approximate the posterior distribution over parameters.
+
+Figure X displays inferred subject params
+
+**NEED SOME FIGURES ON THE INFERRED PARAMS**
+
+### Posterior prediction and model checking
+
+The posterior predictive distribution describes the data we should *expect to see* sampling from our fitted model.
+Intuitively if these predictions do not match the *actual* data used to fit the model, then we have a poor model.
+Recall that in the first step of our Bayesian data analysis we parameters for each subject including $alpha$ the rationality parameter, $\gamma$ their recursive depth, $\lambda$ the cost and referent condition and finally $\sigma$, the degree to which the participant was paying attention during the experiment.
+Figure X displays these parameters.\par
+
+Having inferred these subject-level parameters we’d like to ``run” our RSA-participants back through the experiment.
+That is, for each participant and for each trial they saw (including the object they needed to refer to in that trial) we sample an utterance.
+For example, for a participant who was in the *combined* condition we would sample an utterance given that the actual participant viewed some object on trial 1, repeating the same process for all 32 trials, 24 of which would require the model to refer to a frequent object and 8 requiring the model to refer to an infrequent object.
+Having simulated datasets for each participant we can compare the number of times the RSA-agent and actual participant used the short-form in each condition.
+Correlating model posterior predictive values to human data find close fit $r^2=0.987$.
+Figure X compares these values.
+
+As a baseline we can compare an “optimal Zipfian model”.
+Under this model a speaker in the *combined* condition should always map the short form the most frequent item, but not the infrequent item; in the *time* condition the speaker should always use the short form, regardless of the referent; in the *accuracy* condition the speaker should always use the long forms.
+For comparison, model fit under this model achieves an overall $r^2=0.630$.
+
+```{r plot-inferred-params, fig.env = "figure*", fig.pos = "h", fig.width=6, fig.height=3, fig.align = "center", set.cap.width=T, num.cols.cap=2, fig.cap = "This image spans both columns.
+And the caption text is limited to 0.8 of the width of the document."}
+img <- png::readPNG("figs/inferred-params-plot.png")
+grid::grid.raster(img)
+```
+
 # Results
+
 ```{r plot-post-predictives, fig.env = "figure*", fig.pos = "h", fig.width=6, fig.height=3, fig.align = "center", set.cap.width=T, num.cols.cap=2, fig.cap = "This image spans both columns.
 And the caption text is limited to 0.8 of the width of the document."}
-img <- png::readPNG("figs/post-predictives-plot.png")
+img <- png::readPNG("figs/bda-vs-baseline-plot.png")
 grid::grid.raster(img)
 ```
+
+### Limitations of the current analysis
+There are aspects our current analysis does not capture -- the temporal effects identified by the authors -- however, this remains to be a possible extension of the framework we introduce here.
+Moreover we frame our results in light  argue that this is largely in line with ideas put forth by the authors -- that pragmatic language use may be an essential precursor to more system-level language change.
+In their own words, "pragmatics-driven asymmetry in usage may or may not lead to an immediate shift in lexical representations, it may be an important first step in such a change."\par
 
@@
 Inferring subject-level parameters we are able to fit Kanwal et al.
-(2017) experimental data with high-levels of accuracy ($r^2=0.986).
+(2017) experimental data with high-levels of accuracy ($r^2=0.986$).
 While we believe this provides evidence for the pragmatic-language use interpretation we also highlight the compatibility of this interpretation with an additional insight of Kanwal et al.

### 2019-01-08 paper/paper.Rmd
additional tweaks before ending work today

@@
 ---
-title: "A pragmatic account of Zipf's Law of Abbreviation via Bayesian data analysis"
+title: "Exploring a pragmatic account of Zipf's Law of Abbreviation using Bayesian data analysis"
 bibliography: library.bib
@@
 
-## Rational speech act theory as a model for pragmatic participant behavior
+# Rational speech act theory as a model for pragmatic participant behavior
 
@@
 \begin{equation}
-S_i(u|m) \propto e^{U(u;m)}
+S_i(u|m) \propto e^{-\alpha\times U(u;m)}
 \end{equation}
@@
 \begin{equation}
-U(u;m) = \alpha(log(L_{i-1}(m|u)) - cost(u))
+U(u;m) = -log(L_{i-1}(m|u)) - cost(u)
 \end{equation}
@@
 Defining nested speaker and listener agents could, in principle lead to infinite recursion.
-We take a *literal listener*, denoted $L_0(m|u) as a basecase.
+We take a *literal listener*, denoted $L_0(m|u)$ as a base-case.
 The literal listener does not reason about a speaker model, rather this agent considers the literal semantics of the utterance.
@@
 
-# Modeling framework
+```{r plot-inferred-params, fig.env = "figure*", fig.pos = "h", fig.width=6, fig.height=3, fig.align = "center", set.cap.width=T, num.cols.cap=2, fig.cap = "TEMP - NEEDS UPDATING Inferred subject-level parameter distributions.
+Moving clock-wise from the upper-right facet -- inferred alpha values for subjects, the recursive depth for each subject, the probability that subjects are paying attention during the experiment, the inferred cost / referrent frequency structure used by the subject."}
+img <- png::readPNG("figs/inferred-params-plot.png")
+grid::grid.raster(img)
+```
+
+## Bayesian data analysis with RSA linking function
 
@@
 
-```{r plate-notation, fig.env = "figure", fig.pos = "H", fig.align='center', fig.width=2, fig.height=2, set.cap.width=T, num.cols.cap=1, fig.cap = "One column image."}
+```{r plate-notation, fig.env = "figure", fig.pos = "H", fig.align='center', fig.width=2, fig.height=2, set.cap.width=T, num.cols.cap=1, fig.cap = "THIS IS A PLACE-HOLD AND IS INCORRECT"}
 img <- png::readPNG("figs/plate-temp.png")
@@
 
-For each subject we use $n=200$ samples using Metropolis-Hasting to approximate the posterior distribution over parameters.
-
-Figure X displays inferred subject params
-
-**NEED SOME FIGURES ON THE INFERRED PARAMS**
+For each subject we use $n=200$ samples using Metropolis-Hasting to approximate the posterior distribution over parameters. Figure X displays inferred subject params.
 
@@
 
-```{r plot-inferred-params, fig.env = "figure*", fig.pos = "h", fig.width=6, fig.height=3, fig.align = "center", set.cap.width=T, num.cols.cap=2, fig.cap = "This image spans both columns.
-And the caption text is limited to 0.8 of the width of the document."}
-img <- png::readPNG("figs/inferred-params-plot.png")
-grid::grid.raster(img)
-```
-
 # Results
 
-```{r plot-post-predictives, fig.env = "figure*", fig.pos = "h", fig.width=6, fig.height=3, fig.align = "center", set.cap.width=T, num.cols.cap=2, fig.cap = "This image spans both columns.
-And the caption text is limited to 0.8 of the width of the document."}
+```{r plot-post-predictives, fig.env = "figure*", fig.pos = "h", fig.width=6, fig.height=3, fig.align = "center", set.cap.width=T, num.cols.cap=2, fig.cap = "Posterior predictive values lead to far better fit to human data than a baseline model.
+Horizontal axes display counts of short-form usage from subjects in Kanwal et al.
+(2017).
+Verticle axes display model predictions.
+The left facet displays predicted short-form usage under a baseline model that implements the 'optimal' solution for each condition.
+The right facet displays predicted short-form usage under our BDA model."}
 img <- png::readPNG("figs/bda-vs-baseline-plot.png")

### 2019-01-08 bda/analysis/fit_model.Rmd
minor plot labeling

@@
   gather('model', 'pred', c(baseline, estCnt)) %>%
-  mutate(model=ifelse(model=='estCnt', 'BDA', 'Basline')) %>%
+  mutate(model=ifelse(model=='estCnt', 'BDA', 'Baseline')) %>%
   ggplot(aes(x=actualCnt, y=pred, col=condition), alpha=0.4) +

### 2019-01-11 paper/paper.Rmd
more updates, almost filled in all sections. still needs lots of attention

@@
     Zipf (1935) argued that the relation between a word's frequency and length was 
-    a universal property of human languages emerging from the pressure to minimize speaker-listener effort.
-    Subsequent work has provided indirect evidence for a Zipfian analysis of this relation between frequency and length, what Zipf termed
-    the "Law of Abbreviation". In an effort to directly isolate the impact of speaker and listener pressures,
+    a universal property of human languages emerging from the competing pressures to minimize speaker- and listener-effort during communication.
+    More recent work has explored these ideas, providing additional evidence for a Zipfian analysis of this relation between frequency and length, what Zipf termed
+    the "Law of Abbreviation". However, this work, including Zipf's original analysis, has primarily relied on large-scale observational data, 
+    only providing indirect evidence for the causal role speaker- and listener-pressures have on the "Law of Abbreviation". 
+    In an effort to isolate these competiting dynamics experimentally,
     Kanwal et al. (2017) conducted a study using an artificial language learning paradigm. The authors showed that 
-    only when both a speaker and listener pressure were present did languages consistent with the Law of Abbreviation emerge. The authors
-    highlight that while their data is consistent with a Zipfian analysis, it is also consistent with pragmatic langage use account.
-    We consider this latter account conducting a Bayesian data analysis using the data from Kanwal et al. (2017) representing
+    only when both a speaker- and listener-pressures were present did the Law of Abbreviation emerge. The authors
+    highlight that while their data is consistent with a Zipfian analysis, it is also consistent with an alternative, pragmatic language-use account.
+    We consider this latter account, conducting a Bayesian data analysis using the data from Kanwal et al. (2017) representing
     subjects as rational, pragmatic agents using the Rational Speech Act framework. Fitting subject-level parameters
@@
 Among these, Zipf examined the relation between a word's frequency in a corpus and a variety of properties including its rank-frequency, its denotation size, and its length.
-In terms of this last property, what he later termed the "Law of Abbreviation," Zipf claimed that the more frequent words are, the shorter they tend to be.
-Evidence for this relationship can be found across languages, may also be present in animal communication systems  (Ferrer-i-Cancho et, 2013), and is a property of artificial (programming) languages as well.\par
+In terms of this last property, what he called the "Law of Abbreviation," Zipf claimed that the more frequent words are, the shorter they tend to be.
+Evidence for this relationship can be found across languages, may also be present in animal communication systems  (Ferrer-i-Cancho et, 2013), and has also been observed in artificial (programming) languages (CITATION from Kanwal).\par
 
@@
 Under this framing, communication is viewed as a cooperative, joint activity between interlocutors.
-Importantly, there remains a fundamental asymmetry in any communicative act -- what is effortful for a speaker (production) is different than what is effortful for a listener (understanding).
-Zipf suggested that an optimal language from a speaker perspective would consist of a single, low-cost word which was fully ambiguous - referring to any possible referent in the world.
-In using such a language a listener would need to disambiguate every utterance by the speaker.
+However, this joint activity contains a fundamental asymmetry -- what is effortful for a speaker (production) is different from what is effortful for a listener (understanding).
+An optimal speaker language, Zipf argued, would consist of a single, low-cost word which was fully ambiguous; referring to all possible objects in the world.
+Adopting such a language, a listener would need to disambiguate every utterance by the speaker.
 By contrast, if a language were optimised solely in terms of listener effort, the language should bijectively map all forms to meanings so there was no need for a listener to disambiguate.\par
@@
 Zipf argued that these competing forces, what he called “Speaker’s economy” and “Auditor’s economy”, give rise to the particular distributional forms found in all languages, including the relationship between word frequency and length.
-While this framework is compelling and can largely be seen as foundational to a family of theories examining efficiency in language-structure (CITATIONS) and  use (CITATIONS), the issue is far from settled.
-Subsequent work has argued that properties such as the relation between word frequency and length could also be explained by a process of random typing (Ferrer-i-Cancho & Mascoso del Pardo, 2012).
-Moreover, most evidence supporting the Zipfian approach has been indirect and did not fully determine the causal role of competing speaker and listener pressures.\par
+While this framework is compelling and can largely be seen as foundational to a family of theories examining efficiency in language-structure (CITATIONS) and  -use (CITATIONS), the issue is far from settled.
+Subsequent work has argued that properties such as the relation between word frequency and length can theoretically be explained by a process of random typing (Ferrer-i-Cancho & Mascoso del Pardo, 2012).
+Kanwal et al.
+(2017) highlight the fact that Zipf’s original work, as well as more recent studies analysing large-scale, observational corpora, “do not explicitly target the hypothesized role of communicative pressures…” \par
 
-Kanwal et al.
-(2017) implemented an artificial language learning paradigm to isolate speaker-listener pressures directly.
-In line with Zipf's original analysis, they hypothesized that only when *both* speaker and listener pressures are present in a communication setting would the "Law of Abbreviation" emerge.
-The authors adopted a reference game setting in which participants learned a small language consisting of three words, which could be used to refer to two items.
-They hypothesized that when given a choice between words subjects should prefer to use the less costly, but ambiguous term to refer to a more frequently occurring object.
+To gain traction on the question of the causal role of speaker-listener pressures, Kanwal et al.
+(2017) implemented an artificial language learning paradigm to isolate the pressures directly.
+In line with Zipf's original analysis, they hypothesized that only when *both* speaker- and listener- pressures are present in a communication setting would the "Law of Abbreviation" emerge.
+The authors adopted a miniature artificial language learning set-up in which participants learned a language consisting of three words, which could be used to refer to two items.
+Crucially, two of the words were longer (seven versus three characters) than a third word.
+The two longer words could only be used to refer to one of the two objects.
+The third word, an abbreviation of the longer forms, could be used to refer to either of the two objects (it was ambiguous).\par
+
+Operationalizing production cost as a function to word length, they hypothesized that when given a choice between alternatives, subjects should prefer to use the shorter, less costly term to refer to a more frequently occurring object.
 Importantly, however, this choice should only appear when there was both a speaker- and listener-pressures present.
-Results largely confirmed their hypotheses.\par
+Results indicated that speakers did tend to use the less costly form for the more frequent item when both pressures were present, but, importantly, not in conditions where only one or neither of the pressures were present.\par
 
@@
 That is, it attempts to explain why particular word forms are mapped to particular meanings.
-However, their results are, in theory, consistent with a pragmatic language-use account.\par
+However, their results are, in theory, consistent with a pragmatic language-use account.
+We refer the reader to their paper for a more detailed discussion, but include the essential elements here:\newline
 
-*"There is a distinction between a language-user's mental representation of the lexicon, and the form-meaning mappings they actually produce in communication...the nature of the communication task in this experiment may have caused them to produce only the short form for one object and the long form for the other based on purely pragmatic considerations."*\par
+*"There is a distinction between a language-user's mental representation of the lexicon, and the form-meaning mappings they actually produce in communication...the nature of the communication task in this experiment may have caused them to produce only the short form for one object and the long form for the other based on purely pragmatic considerations."*\newline
 
-Put differently, the underlying lexicon learned by subjects may have remained unchanged during the experiment -- a finding which, on its face, is at odds with the claim that this experiment demonstrates lexicon-level change.
+In other words, the underlying lexicon learned by subjects may have remained unchanged during the experiment -- a finding which, on its face, is at odds with the claim that this experiment demonstrates lexicon-level change.
 Rather, subjects may have *used* the lexicon pragmatically, preferring to map less costly forms to more frequent objects, while leaving the underlying lexicon unchanged.
-The authors’ correctly point out that because they only recorded participants’ actual language production they cannot distinguishing between the lexicon change account and pragmatic accounts.
+The authors’ highlight that because they only recorded participants’ actual language production they cannot distinguish between the lexicon-change account and pragmatic, language-use account.\par
 
@@
 To preview our results we find good fit from model posterior-predictive values to Kanwal et al.
-(2017) participant data ($R^2=0.986$), over and above what would would be expected from a baseline model ($r^2=0.68$).
+(2017) participant data ($r^2=0.987$), over and above what would would be expected from a baseline model ($r^2=0.684$).
 We believe this provides compelling evidence in support of the alternative hypothesis outlined by Kanwal et al.
@@
 (2017), their hypotheses and the alternative interpretation of their results.
-We then introduce our general approach to the Bayesian data analysis as well as the Rational Speech-act framework we will use to model pragmatic language use.
-We evaluate our results summarising the findings and highlighting aspects that our modeling does not capture.
+We then introduce our general approach to the Bayesian data analysis as well as the Rational Speech-act framework we will use to model subject’s pragmatic language use.
+We evaluate our results summarising the findings and highlighting dimensions of their analysis our current implementation does not yet capture.
 Following this work we return the question posed by Kanwal et al.
@@
 
+```{r plot-inferred-params, fig.env = "figure*", fig.pos = "h", fig.width=6, fig.height=3, fig.align = "center", set.cap.width=T, num.cols.cap=2, fig.cap = "TEMP - NEEDS UPDATING Inferred subject-level parameter distributions.
+Moving clock-wise from the upper-right facet -- inferred alpha values for subjects, the recursive depth for each subject, the probability that subjects are paying attention during the experiment, the inferred cost / referrent frequency structure used by the subject."}
+img <- png::readPNG("figs/inferred-params-plot.png")
+grid::grid.raster(img)
+```
+
 # Experimental evidence of Zipf's Law of Abbreviation
 
-Kanwal et al.
-(2017) used an artificial language learning paradigm to isolate the impact of speaker- and listener-pressures on language structure.
-During training, participants first learn a miniature language.
-During test, participant’s language use is recorded, typically over a series of communication events or across multiple generations of participants.
-This paradigm has been used to study changes in language structure (CITATIONS).\par
+Artificial language learning paradigms have proven to be a productive methodology for studying language change in the laboratory (or online via web experiments) (CITATIONS).
+In the typical set-up, participants are trained on a simple language -- a series of form-meaning mappings, often using nonce words with novel objects.
+During test, participant’s language producing is recorded, typically over a series of communication events or across multiple generations of participants accomplishing some task (CITATIONS).
+Subsequent analyses focus on how properties of the produced language change as it used over time participants.\par
 
 Kanwal et al.
-(2017) adopt this basic framework -- participants learned a miniature artificial language and were asked to communicate with an interlocutor in a reference game setting.
+(2017) adopt this basic framework -- participants learned a miniature artificial language and were subsequently asked to communicate with an interlocutor in a reference game setting.
 Following ideas first indicated by Zipf (1935), the authors hypothesized the presence or absence of pressures to communicate quickly (speaker pressure) and accurately (listener pressure)  should impact the resulting languages.
-The study proceeded as follows.
-Participants were assigned to one of four possible conditions.
-A *combined* condition in which both speaker- and listener-pressures were present.
-An *accuracy* condition, in which only a listener pressure was present.
-A *time* condition in which only a speaker pressure was present.
-They also tested a *neither* condition in which neither speaker nor listener pressures were present.
-For the purpose of our current project, we do not model the *neither* condition as it's unclear what optimal behavior should like.\par
+Participants were assigned to one of four possible conditions which varied in the pressures that were present.
+Both speaker- and listener-pressures were present in the *combined* condition, only the listener pressure present in the *accuracy* condition, only the speaker pressure present in the *time* conding.
+They also tested a fourth *neither* condition in which neither speaker nor listener pressures were present.
+For the purpose of our current project, we do not report our results on the control, *neither* condition as behavior should be arbitrary.\par
 
-For a comprehensive description of the experimental set-up we refer the reader to Kanwal et al.
+For a comprehensive description of the experimental methodology we refer the reader to Kanwal et al.
 (2017), focusing on essential details in this section.
 During training participants learned labels for two novel objects.
-Importantly, one of the objects occurred more frequently than the other.
+Importantly, one of the objects occurred three times as frequently than the other (appearing 24 vs 8 times).
 During training participants observed the objects at their relative frequencies as well as labels.
@@
 Importantly, in this set-up “zop” was ambiguous between the two object referents.
-During test, pairs of participants were asked to participate in a communication game in which a “director” was shown an object” and asked to transmit its name to the “matcher.” The “director” was always given a choice to use the long-form label or the ambiguous short-form.
+During test, participants were asked to participate in a communication game in which a “director” was shown an object and asked to transmit its label to the “matcher.” The “director” was always given a choice to use the long-form label or the ambiguous short-form.
 Crucially, to transmit the label to the matcher the director had to click on it, holding the mouse while each character appeared sequentially.
 In this setting, transmitting a longer-form required more time -- an elegant operationalization of speaker production cost.
-Listener pressure was operationalized via a pressure for the matcher to correctly identify the referent.\par
+Listener cost was operationalized via a pressure for the matcher to correctly identify the referent.\par
 
-The authors hypothesized that a preference to map the short form to the more frequent object, but not the infrequent object should only occur in *combined* condition.
-By contrast, when there was no pressure to communicate (no listener was present in the *time* condition), and only the speaker cost was present, participants should always prefer the shortened form, regardless of the referent.
-When there was no cost difference between the shortened and lengthened forms, but a listener was present, then participants should always prefer to use the unambiguous, long-forms regardless of referentin the *accuracy* condition.
-Results were largely consistent with these hypotheses.
+The authors hypothesized that a preference to map the short form to the frequent object, but not the infrequent object, should emerge in the *combined* condition.
+Recall that in this condition  there was both a pressure to *communicate quickly* (subjects were incentivized based on their time) and accurately (subjects were incentivized based on their matching performance).
+By contrast, in the *time* condition, in which there was no pressure to communicate accurately (no listener was present), and only the speaker cost was present, participants should always prefer the shortened form, regardless of the referent.
+In the *accuracy* condition there was no cost difference between the shortened and lengthened forms, but a listener was present.
+In this condition, participants should always prefer to use the unambiguous, long-form labels to maximize matching performance.
+Results across all three conditions were largely consistent with these predictions -- only in the *combined* speaker-listener pressure condition did the Law of Abbreviation emerge via shortening of the label for the most frequent object, but not the infrequent object.
 
@@
 
-While the data is consistent with a Zipfian account of the Law of Abbreviation, Kanwal et al.
-(2017) highlight an important alternative explanation.
+While these results are consistent with a Zipfian account of the Law of Abbreviation, Kanwal et al.
+(2017) highlight an alternative explanation.
 They describe a distinction between the "mental representation" of the lexicon and produced forms.
 That is, Zipf's Law of Abbreviation is an account of lexical change, not language-use.
-If participants are retaining both words in their "mental lexicon," but *using the lexicon pragmatically* by using the shortened form only for the more frequent item, then the lexicon itself has not "changed" at all.
-This analysis falls largely in line with Gricean notion of natural language pragmatics.\par
+If participants are retaining both the long-forms and short-form word in their "mental lexicon," but *using the lexicon pragmatically*, then the lexicon itself has not "changed." This analysis falls largely in line with Gricean notion of natural language pragmatics.\par
 
-Grice (1957) argued that communication could largely be analysed as an instance of rational, cooperative behavior.
+Grice (1975) argued that communication could largely be understood as an instance of rational, cooperative behavior.
 Under this framework speakers choose utterances to convey meanings.
-Listeners attempt to infer the speaker's communicative goal.
+Listeners attempt to infer the speaker's intended meaning given utterances.
 At the heart of Grice’s theory was a set of conversational maxims known to both speakers and listeners (be truthful, relevant, informative and perspicuous).
-Inferences about a speaker’s intended meaning could be derived from these maxims by a listener.
-In particular, it is assumed that the listener can reason about the speaker and information that is mutually known to both.
-The matcher can use such information, such as the fact that using the short-form is less costly than using a long-form, and that the frequent object occurs three times more often than the infrequent object.
-The matcher and director might reason as follows: *If the director is trying to transmit information as quickly as possible they should always use the quicker, shortened form.
-But if they want me to pick out the correct referent they should always transmit the slower, longer form.
-An optimal solution should use the short form in in as many trials as possible.
-Therefore we should map the shorter, ambiguous form to the more frequent object.*\par
+Inferences about a speaker’s intended meaning could be derived from speaker behavior relative to these maxims.
+In particular, this framework assumes that the listener can reason about the speaker’s intentions and relevant contextual information (e.g.
+common ground).\par
 
-Notably, by reasoning about one-another, about the goals of the game (be quick, be accurate), and the details of context (differential word costs and object frequencies) a pair can converge on optimal *use* of the provided language, but *without changing the underlying structure*.
+Translating this framing to the current project, we assume that the “matcher” can use contextual information.
+For example, in the *combined* condition the matcher knows that it is mutually beneficial for them to complete the task as quickly and as accurately as possible.
+They know the short label is faster (less costly to the speaker) than using a long-form, and that the frequent object occurs three times more often than the infrequent object.
+In a given trial, in which the matcher is trying to complete the task a Gricean description of their reasoning might go as follows: \newline
+*If the director is trying to transmit information as quickly as possible they should always use the faster, shortened form.
+But if they want me to pick out the correct referent they should always transmit the unambiguous, slower form.
+An optimal trade-off should use the short form in as many trials as possible without incurring undo ambiguity.
+Therefore we should map the shorter, ambiguous form to the more frequent object.*\newline
+
+Notably, by reasoning about one-another, about the goals of the game (be quick, be accurate), and the details of context (differential word costs and object frequencies) a pair can converge on optimal *use* of the provided language *without changing the underlying structure*.
 
@@
 RSA has proven to be a productive framework for modeling a range of pragmatic phenomena from hyperbole, metaphor, scalar implicature and others (CITATIONS).
-Under RSA a speaker agent describes a conditional distribution mapping meanings $u\in U$ to utterances $m\in M$, written as $S(u|m)$.
+Under RSA a speaker agent describes a conditional distribution, mapping meanings $u\in U$ to utterances $m\in M$, written as $S(u|m)$.
 A speaker agent describes a conditional distribution mapping from utterances to meaning, written as $L(m|u)$.
@@
 \end{equation}
-Defining nested speaker and listener agents could, in principle lead to infinite recursion.
-We take a *literal listener*, denoted $L_0(m|u)$ as a base-case.
+Defining nested speaker and listener agents could, in principle, lead to infinite regress.
+We take a *literal listener*, denoted $L_0(m|u)$, as a base-case.
 The literal listener does not reason about a speaker model, rather this agent considers the literal semantics of the utterance.
@@
 
-```{r plot-inferred-params, fig.env = "figure*", fig.pos = "h", fig.width=6, fig.height=3, fig.align = "center", set.cap.width=T, num.cols.cap=2, fig.cap = "TEMP - NEEDS UPDATING Inferred subject-level parameter distributions.
-Moving clock-wise from the upper-right facet -- inferred alpha values for subjects, the recursive depth for each subject, the probability that subjects are paying attention during the experiment, the inferred cost / referrent frequency structure used by the subject."}
-img <- png::readPNG("figs/inferred-params-plot.png")
+## Bayesian data analysis with RSA linking function
+
+To assess the degree to which the results in Kanwal et al.
+(2017) can be described by a pragmatic, language-use account, we conduct a Bayesian data analysis, using RSA as our linking function.
+Our analysis proceeds in two parts.
+In the *parameter inference* phase we infer subject-level parameters.
+In the *model-checking and posterior prediction* phase we assess how well our model, with the inferred parameters, describes the experimental data.
+
+### Parameter inference
+
+```{r table-parameters, fig.env = "table", fig.pos = "H", fig.align='center', fig.width=3.2, fig.height=2, set.cap.width=T, num.cols.cap=1, fig.cap = "Parameters"}
+img <- png::readPNG("figs/parameters-table.png")
 grid::grid.raster(img)
@@
 
-## Bayesian data analysis with RSA linking function
-
-Top assess the degree to which the observed production data in Kanwal et al.
-(2017) can be described by a pragmatic language-use account we conduct a Bayesian data analysis, using RSA as our linking function.
-Our analysis proceeds in two parts.
-In the parameter inference phase we infer subject-level parameters.
-In the model-checking and posterior prediction phase we assess how well our model, with the inferred parameters, describes the experimental data.
-
-### Parameter inference
-
-For each participant we infer a set of parameters -- $\alpha$, $\beta$ are RSA model parameters corresponding the the subjects rationality and recursive depth, respectively.
-The parameters $\gamma$ and $\lambda$ correspond to the particular condition (made up of utterance costs and object referent priors) as wells as a focus parameter describing the probability that the subject is not paying attention and choosing to produce utterance randomly.
+We model each participant as an RSA agent, inferring for each the following set of parameters -- $\alpha$, $\beta$ are RSA model parameters corresponding the the subjects rationality and recursive depth, respectively.
+The parameters $\gamma$ and $\lambda$ correspond to experiment-specific parameters -- the particular condition (described by the utterance costs and object referent priors), and a focus parameter describing the probability that the subject is not paying attention and choosing to produce utterance randomly.
 Table 1 includes the parameters and their descriptions.\par
@@
 ```
-
-For each subject we use $n=200$ samples using Metropolis-Hasting to approximate the posterior distribution over parameters. Figure X displays inferred subject params.
+For each subject we use $n=200$ samples using Metropolis-Hasting to approximate the posterior distribution over parameters. Figure 1. displays inferred subject params across the all participants.
 
@@
 Intuitively if these predictions do not match the *actual* data used to fit the model, then we have a poor model.
-Recall that in the first step of our Bayesian data analysis we parameters for each subject including $alpha$ the rationality parameter, $\gamma$ their recursive depth, $\lambda$ the cost and referent condition and finally $\sigma$, the degree to which the participant was paying attention during the experiment.
-Figure X displays these parameters.\par
-
-Having inferred these subject-level parameters we’d like to ``run” our RSA-participants back through the experiment.
-That is, for each participant and for each trial they saw (including the object they needed to refer to in that trial) we sample an utterance.
-For example, for a participant who was in the *combined* condition we would sample an utterance given that the actual participant viewed some object on trial 1, repeating the same process for all 32 trials, 24 of which would require the model to refer to a frequent object and 8 requiring the model to refer to an infrequent object.
-Having simulated datasets for each participant we can compare the number of times the RSA-agent and actual participant used the short-form in each condition.
-Correlating model posterior predictive values to human data find close fit $r^2=0.987$.
-Figure X compares these values.
-
-As a baseline we can compare an “optimal Zipfian model”.
-Under this model a speaker in the *combined* condition should always map the short form the most frequent item, but not the infrequent item; in the *time* condition the speaker should always use the short form, regardless of the referent; in the *accuracy* condition the speaker should always use the long forms.
-For comparison, model fit under this model achieves an overall $r^2=0.630$.
-
-# Results
+Recall that in the first step of our Bayesian data analysis we infer parameters for each subject including the rationality parameter ($\alpha$), their recursive depth ($\beta$),  the cost and referent condition structure ($\gamma$), and their level of attention ($\lambda$).\par
 
@@
 (2017).
-Verticle axes display model predictions.
+Vertical axes display model predictions.
 The left facet displays predicted short-form usage under a baseline model that implements the 'optimal' solution for each condition.
@@
 
+Having inferred these subject-level parameters we’d like to ``run” our RSA-participants back through the experiment.
+That is, for each participant and for each trial they saw, we sample an utterance (e.g.
+$u \sim S_{\text{subject}_i,\text{ trial}_j}(u|m)$).
+For example, for a participant in the *combined* condition we would sample an utterance given that the actual participant viewed some object on trial 1, repeating the same process for all 32 trials, 24 of which would require the model to refer to a frequent object and 8 requiring the model to refer to an infrequent object.
+Having simulated datasets for each participant we can compare the number of times the RSA-agent and actual participant used the short-form in each condition following the analysis in Kanwal et al.
+(2017).
+
+#  Results
+
+As a baseline we can compare a deterministic speaker using the optimal Zipfian language in each condition.
+Under this model a speaker in the *combined* condition should always map the short form the most frequent item, but not the infrequent item; in the *time* condition the speaker should always use the short form, regardless of the referent; in the *accuracy* condition the speaker should always use the long forms.
+This baseline describes a speaker who deterministically chooses an utterance based on the optimal strategy.
+
+To assess model performance we correlate model posterior predictive values for our RSA-subjects and baseline mode to human data.
+Table 2 displays $r^2$ values for both models.
+We find close fit for our rational pragmatic agents, significantly above our baseline model.
+Figure X shows the values.
+
+\begin{tabular}{ |p{2cm}|p{1cm}|  }
+ \hline
+  Model & $r^2$ \\
+ \hline
+ Baseline & $0.630$ \\
+ BDA RSA & $0.987$ \\
+ \hline
+\end{tabular}
+
 ### Limitations of the current analysis
-There are aspects our current analysis does not capture -- the temporal effects identified by the authors -- however, this remains to be a possible extension of the framework we introduce here.
+While our current framework accounts for a large percentage of variance in the human data, there are aspects our current analysis does not capture.
+Figure 5.
+Of Kanwal et al.
+(2017) shows temporal effects throughout the experiment -- speakers are more likely to use the shorter form later in the experiment compared to earlier in the experiment.
+Our current model does not capture this particular effect.
+However, there are various ways we might try to instantiate temporal effects corresponding to this kind of convergence.
+In our current framework this might be instantiated via reduction in uncertainty over the prior frequency distributions for the objects referents.
+That is, uncertainty at the social-epistemic level -- does your partner have the same priors you do.
+Taking another angle we might assume that the cost of the long-form becomes more onerous as the task proceeds, making speakers even more likely to want to avoid it.
+Our current framework provides many opportunities for this type of follow-up.\par
+
+  -- the temporal effects identified by the authors -- however, this remains to be a possible extension of the framework we introduce here.
 Moreover we frame our results in light  argue that this is largely in line with ideas put forth by the authors -- that pragmatic language use may be an essential precursor to more system-level language change.

### 2019-01-11 bda/analysis/fit_model.Rmd
update human-model comparison to be proportions instead of counts (to mirror Kanwals analysis). think about t statistical test for the correlation coefs?

--- 2019-01-07 bda/analysis/fit_model.Rmd
@@
 
+## Counts plot 1
+
 ```{r}
@@
 
+## Counts plot 2
+
 Compare to "optimal baseline" optimal in that `time` condition always pick short, `accuracy` always long, etc.
@@
   mutate(baseline=ifelse(condition=='combined' & type=='frequent', 24,
-                         ifelse(condition=='combined' & type=='infrequenty', 8,
+                         ifelse(condition=='combined' & type=='infrequent', 8,
                                 ifelse(condition=='accuracy' & type=='frequent', 0,
@@
   gather('model', 'pred', c(baseline, estCnt)) %>%
-  mutate(model=ifelse(model=='estCnt', 'BDA', 'Basline')) %>%
+  mutate(model=ifelse(model=='estCnt', 'BDA', 'Baseline')) %>%
   ggplot(aes(x=actualCnt, y=pred, col=condition), alpha=0.4) +
-    geom_point(aes(fill=condition), alpha=0.6, size=3, colour="white", pch=21) +
+    geom_point(aes(fill=condition, shape=type), alpha=0.6, size=3) +
     geom_abline(intercept=0, slope=1, col='grey', alpha=0.4) +
@@
 
+# Proportions Analysis
+
 ```{r}
-df_predictives2 %>%
+df_cors <- df_predictives2 %>%
   mutate(baseline=ifelse(condition=='combined' & type=='frequent', 24,
-                         ifelse(condition=='combined' & type=='infrequenty', 8,
+                         ifelse(condition=='combined' & type=='infrequent', 8,
+                                ifelse(condition=='accuracy' & type=='frequent', 0,
+                                       ifelse(condition=='accuracy' & type=='infrequent', 0,
+                                              ifelse(condition=='time' & type=='frequent', 48, 16)))))) %>%
+  gather('model', 'pred', c(baseline, estCnt)) %>%
+  group_by(IP, condition, actualCnt, model) %>%
+  summarise(cntShortForms=sum(pred)) %>%
+  group_by(IP, model, condition) %>%
+  summarise(estCnts=sum(cntShortForms),
+            actCnts=sum(actualCnt)) %>%
+  mutate(condition_n=ifelse(condition=='combined' | condition=='accuracy', 32, 64)) %>%
+  mutate(estProp=estCnts/condition_n,
+         actProp=actCnts/condition_n) %>%
+  ungroup
+```
+
+```{r}
+df_cors %>%
+  group_by(model) %>%
+  summarise(n=n(),
+            r=cor(estProp, actProp)^2)
+```
+
+## Proportions plot1
+
+```{r}
+df_cors %>%
+  mutate(model=ifelse(model=='baseline', 'Baseline', 'BDA RSA')) %>%
+  ggplot(aes(x=actProp, y=estProp, col=condition), alpha=0.4) +
+    geom_point(aes(fill=condition), alpha=0.6, size=3) +
+    geom_abline(intercept=0, slope=1, col='grey', alpha=0.4) +
+    # geom_point(aes(x=actualInrequentProps, y=estInfrequentProps, col=condition), alpha=0.4) +
+    facet_wrap(~model) +
+    ylab('Baseline model \n counts of short form') +
+    xlab('Human data \n counts of short form') +
+    ylim(0, 1) +
+    xlim(0, 1) +
+    theme_few()
+
+ggsave('../../paper/figs/bda-vs-baseline-plot.png', plot = last_plot(), width=6, height=3)
+```
+
+Some kind of significance test?
+
+```{r}
+fisher.z <- function(r1,r2,n1,n2) {
+  # Todo: review (https://stats.stackexchange.com/questions/64152/are-two-pearson-correlation-coefficients-different)
+  ((0.5*log((1+r1)/(1-r1)))-(0.5*log((1+r2)/(1-r2))))/((1/(n1-3))+(1/(n2-3)))^0.5
+}
+
+```
+
+```{r}
+rs <- df_predictives2 %>%
+  mutate(baseline=ifelse(condition=='combined' & type=='frequent', 24,
+                         ifelse(condition=='combined' & type=='infrequent', 8,
                                 ifelse(condition=='accuracy' & type=='frequent', 0,
@@
                                               ifelse(condition=='time' & type=='frequent', 48, 16)))))) %>%
-  group_by(condition) %>%
-  summarise(r2 = cor(actualCnt, estCnt)^2)
-
-df_predictives2 %>%
-  mutate(baseline=ifelse(condition=='combined' & type=='frequent', 24,
-                         ifelse(condition=='combined' & type=='infrequenty', 8,
-                                ifelse(condition=='accuracy' & type=='frequent', 0,
-                                       ifelse(condition=='accuracy' & type=='infrequent', 24,
-                                              ifelse(condition=='time' & type=='frequent', 48, 16)))))) %>%
-  summarise(r2 = cor(actualCnt, baseline))
-
+  filter(type=='frequent') %>%
+  gather(model, value,  c(estCnt, baseline)) %>%
+  select(actualCnt, model, value, condition) %>%
+  group_by(model) %>%
+  summarise(n=n(),
+            r=cor(actualCnt, value)^2)
+print(rs)
 ```

### 2019-01-11 bda/analysis/fit_model.Rmd
updated parameter sections, initial general discussion

@@
   mutate(baseline=ifelse(condition=='combined' & type=='frequent', 24,
-                         ifelse(condition=='combined' & type=='infrequent', 8,
+                         ifelse(condition=='combined' & type=='infrequent', 0,
                                 ifelse(condition=='accuracy' & type=='frequent', 0,

### 2019-01-11 paper/paper.Rmd
update human-model comparison to be proportions instead of counts (to mirror Kanwals analysis). think about t statistical test for the correlation coefs?
updated parameter sections, initial general discussion

@@
     \author{{\large \bf Benjamin N. Peloquin} \\ \texttt{bpeloqui@stanford.edu} \\ Department of Psychology \\ Stanford University
-    \And
-    {\large \bf Noah D. Goodman} \\ \texttt{noah@university.edu} \\ Department of Computer Science \\ Stanford University
+    \And {\large \bf Joseph Cornelius} \\ \texttt{jocorn@stanford.edu} \\ Department of Psychology \\ Stanford University
+    \And {\large \bf Noah D. Goodman} \\ \texttt{noah@university.edu} \\ Department of Computer Science \\ Stanford University
     \And {\large \bf Michael C. Frank} \\ \texttt{mike@university.edu} \\ Department of Psychology \\ Stanford University}
@@
 In a given trial, in which the matcher is trying to complete the task a Gricean description of their reasoning might go as follows: \newline
-*If the director is trying to transmit information as quickly as possible they should always use the faster, shortened form.
+
+\textit{If the director is trying to transmit information as quickly as possible they should always use the faster, shortened form.
 But if they want me to pick out the correct referent they should always transmit the unambiguous, slower form.
 An optimal trade-off should use the short form in as many trials as possible without incurring undo ambiguity.
-Therefore we should map the shorter, ambiguous form to the more frequent object.*\newline
+Therefore we should map the shorter, ambiguous form to the more frequent object.}\newline
 
@@
 In the *parameter inference* phase we infer subject-level parameters.
-In the *model-checking and posterior prediction* phase we assess how well our model, with the inferred parameters, describes the experimental data.
-
-### Parameter inference
-
-```{r table-parameters, fig.env = "table", fig.pos = "H", fig.align='center', fig.width=3.2, fig.height=2, set.cap.width=T, num.cols.cap=1, fig.cap = "Parameters"}
-img <- png::readPNG("figs/parameters-table.png")
-grid::grid.raster(img)
-```
+In the *model-checking and posterior prediction* phase we assess how well our model, with the inferred parameters, describes the experimental data.\par
 
 We model each participant as an RSA agent, inferring for each the following set of parameters -- $\alpha$, $\beta$ are RSA model parameters corresponding the the subjects rationality and recursive depth, respectively.
-The parameters $\gamma$ and $\lambda$ correspond to experiment-specific parameters -- the particular condition (described by the utterance costs and object referent priors), and a focus parameter describing the probability that the subject is not paying attention and choosing to produce utterance randomly.
-Table 1 includes the parameters and their descriptions.\par
+The parameters $\lambda$ and $\gamma$ correspond to experiment-specific parameters.
+We describe each parameter in more depth.
+
+#### Rationality parameter ($\alpha$)
+Rational speech-act theory frames natural language understanding and generation as probabilistic inference.
+At the heart of this model is a utility calculation (equation 2) in which the speaker considers both the informativity of using an utterance $u$ to convey some meaning $m$, as well as it’s cost $cost(u)$.
+We capture the degree to which the speaker has as preference to choose the most informativity utterance by multiplying this utility function by a scalar $\alpha$.
+We infer a separate $\alpha$ value for each subject using a uniform prior over the integers $1...30$.\par
+
+#### Recursive depth ($\beta$)
+Perhaps most characteristic of RSA is the nested model definitions in which speakers and listeners are defined recursively in terms of one another.
+The parameter $\lambda$ controls the depth of recursion.
+We assume a uniform prior over the integers $1..3$ inferring the level of recursion for each subject.\par
+
+#### Cost and meaning structure ($\lambda$)
+Rather than explicitly map each subject to one of the three conditions of we consider in this project (*combined*, *time*, and *accuracy*) we allow the model to infer these from the data.
+For each condition we describe the particular utterance cost and meanings prior structure.
+In the *combined* condition we consider utterance costs such that the longer forms are twice as expensive as shorter forms and meaning priors the reflect the training proportions (the infrequent item occurs 8 times and the frequent items 24 times).
+In the *time condition* we assume the same structure except we only consider an $S_0$ speaker who does not consider a listener agent.
+(Recall that no listener agent is present in that condition.) For the *accuracy* condition we consider the same structure as in the *combined* condition, except we assume that there is no cost difference between the utterances.
+We also consider a fourth condition which we call *other*, which allows us to describe distintinctly suboptimal behavior.
+Some respondents nearly always used the shorter form for the infrequent object.
+This behavior is consistent with reversing the prior object frequencies.
+We use a categorical prior over these cost-meaning structure assuming that the the *other* condition is far less likely than the others.\par
+
+#### Attention during the experiment ($\gamma$)
+Finally, we consider that subjects may be employing a rational strategy, but might be distracted during the experiment.
+Our attention parameter $\lambda$ captures the probability that a participants chooses an utterance randomly.
+We use a $\beta(12, 1)$ prior to encode that we expect participants to be paying attention and unlikely to choose randomly during the experiment.\par
+
+The top-right facet of Figure 1 shows the inferred parameters across all participants.
+In particular, we note that few participants were categorized as using the suboptimal *other* condition strategy and that we inferred that most participants were paying attention throughout the course of the experiment ($\alpha$ levels are well above 0.9).
 
@@
 ```
-For each subject we use $n=200$ samples using Metropolis-Hasting to approximate the posterior distribution over parameters. Figure 1. displays inferred subject params across the all participants.
+We infer  each subject we use $n=200$ samples using Metropolis-Hasting to approximate the posterior distribution over parameters.
+Figure 1.
+displays inferred subject params across the all participants.
 
@@
 ```
-
 Having inferred these subject-level parameters we’d like to ``run” our RSA-participants back through the experiment.
-That is, for each participant and for each trial they saw, we sample an utterance (e.g.
-$u \sim S_{\text{subject}_i,\text{ trial}_j}(u|m)$).
-For example, for a participant in the *combined* condition we would sample an utterance given that the actual participant viewed some object on trial 1, repeating the same process for all 32 trials, 24 of which would require the model to refer to a frequent object and 8 requiring the model to refer to an infrequent object.
-Having simulated datasets for each participant we can compare the number of times the RSA-agent and actual participant used the short-form in each condition following the analysis in Kanwal et al.
-(2017).
+Intuitively this means that we consider the actual sequence of trials for each subject.
+For a particular trial we examine the particular object they are asked to refer to and pass this two our RSA speaker model with that subject’s inferred parameters, generating a corresponding utterance (i.e.
+we sample $u \sim S_{\text{subject}_i,\text{ trial}_j}(u|m)$).
+In this way we can compare trial-by-trial production data of the RSA agents to actual subject data.
 
@@
 
-As a baseline we can compare a deterministic speaker using the optimal Zipfian language in each condition.
+Following the analysis in Kanwal et al.
+(2017) we examine the proportion of trials in which a particular subject used the short-form (see Kanwal Figure 5).
+That is, we compare the proportion of trials in which subject $i$ used the short form, comparing this to proportion of trials predicted by our RSA model and by a baseline comparison.
+
+### Baseline model
+As a baseline we compare a deterministic speaker using the optimal Zipfian language in each condition.
 Under this model a speaker in the *combined* condition should always map the short form the most frequent item, but not the infrequent item; in the *time* condition the speaker should always use the short form, regardless of the referent; in the *accuracy* condition the speaker should always use the long forms.
@@
 Table 2 displays $r^2$ values for both models.
-We find close fit for our rational pragmatic agents, significantly above our baseline model.
-Figure X shows the values.
+We find close fit for our rational pragmatic agents, significantly above our baseline model
 
@@
  \hline
- Baseline & $0.630$ \\
- BDA RSA & $0.987$ \\
+ Baseline & $0.466$ \\
+ BDA RSA & $0.972$ \\
  \hline
 \end{tabular}
+
+# General discussion
+
+Zipf (1935) argued that human behavior could largely be understood in terms of effort minimization.
+In the domain of language, he argued that the particular distributional properties we find in large-scale corpora, such as the fact that more frequent words tend to be shorter, provide evidence for his effort-minimization framework.
+Kanwal et al.
+(2017) presented an experimental test of the Zipfian hypothesis conducting a miniature language-learning experiment and isolating the impact a speaker- and listener-pressures directly.
+In their analysis Kanwal et al.
+(2017) highlighted that their results were consistent with a Zipfian, language-change account as well as a pragmatic langauge-use account.
+This distinction, the argued, cannot be assessed with the current set of results as the data collected only constitutes production data, there is not information about the underlying lexicon's the subjects are using.par
+
+To test the pragmatic, language-use account we conducted a Bayesian data analysis, modeling subjects as rational pragmatic agents using the Rational Speech Act framework (Frank & Goodman, 2012; Goodman & Frank, 2016).
+Inferring subject-level parameters we find good fit with the empirical data providing evidence that subject behavior in this experiment may a result of pragmatic langauge use, rather than lexicon change.\par
 
@@
 However, there are various ways we might try to instantiate temporal effects corresponding to this kind of convergence.
-In our current framework this might be instantiated via reduction in uncertainty over the prior frequency distributions for the objects referents.
+For example, this might be instantiated via reduction in uncertainty over the prior frequency distributions for the objects referents.
 That is, uncertainty at the social-epistemic level -- does your partner have the same priors you do.
 Taking another angle we might assume that the cost of the long-form becomes more onerous as the task proceeds, making speakers even more likely to want to avoid it.
-Our current framework provides many opportunities for this type of follow-up.\par
+Our current framework provides many opportunities for this type of follow-up and we consider this as an important area for further investigation\par
 
-  -- the temporal effects identified by the authors -- however, this remains to be a possible extension of the framework we introduce here.
-Moreover we frame our results in light  argue that this is largely in line with ideas put forth by the authors -- that pragmatic language use may be an essential precursor to more system-level language change.
-In their own words, "pragmatics-driven asymmetry in usage may or may not lead to an immediate shift in lexical representations, it may be an important first step in such a change."\par
-
-# Connecting pragmatic language use to language change
-
-# Conclusion
-
-Languages around the world display a consistent relationship between word length and frequency -- more frequent words tend to be shorter (CITATIONS).
-Zipf (1935) called this particular finding the “Law of Abbreviation” and characterized it in terms of competing speaker listener pressures.
-While other work has provided indirect evidence consistent with this Zipfian analysis (CITATIONS), Kanwal et al.
-(2017) attempted to derive this effect while isolating speaker and listener pressures experimentally.
-In an artificial language learning paradigm the authors showed that subjects were significantly more likely to produce lexicons consistent the “Law of Abbreviation,” but only when both speaker and listener pressures are present.
-While this evidence is compelling the authors highlight an alternative hypothesis that may have generated the same effect, but could not be distinguished in the current study -- pragmatic language use instead of lexical change.
-To test this secondary hypothesis we conducted a Bayesian data analysis, modeling participants as rational pragmatic agents using the Rational Speech-act (RSA) framework.
-Inferring subject-level parameters we are able to fit Kanwal et al.
-(2017) experimental data with high-levels of accuracy ($r^2=0.986$).
-While we believe this provides evidence for the pragmatic-language use interpretation we also highlight the compatibility of this interpretation with an additional insight of Kanwal et al.
-(2017) -- pragmatic language use may serve as an important bootstrap for future language change.
+### The connection between rational langauge use and langauge change
+Recent work has started to draw connection betwen local, in-the-moment language use as modeled by rational, pragmatic agents and aggregate, distribuation-level properties of natural langauge (Goodman & Frank, 2016; Levy, 2018).
+This is not without precedent.
+Zipf (1935) implicitly argues language may be optimized at the level of conversation in his example of an optimal speaker and listener language.
+He describes a set-up in which there are a set of utterance $U$ and meanings $M$ which differ in costs and frequency, respectively.
+While Zipf was not more explicitly in diescribing the level at which language optimization occurs, his examples mirror the reference game setting we have adopted here.
+Morever, the explicit connection with rational, pragmatic langauge use was described by Horn (1984) who argued that Zipf's speaker and listener pressures ("Speaker's economy" and "Auditor's ecnonomy") were fundamental to Grice's (1957) derivation of his Cooperative Principle and associated conversational maxims.
+In general, it is not hard to see a scenario in which pragmatic language use bootstraps later lexicon-level change via processes like inter-generational transmission.
+Kanwal et al.
+(2017) present this very idea writing,"pragmatics-driven asymmetry in usage may or may not lead to an immediate shift in lexical representations, it may be an important first step in such a change." Take the exmaple of the current experiment.
+In the *combined* some subjects produced langauges which only consisted of the shortened form for the frequent object and the long form for the infrequent object.
+If a subsequent generation of speakers were to learn from this data, they would never be exposed to the original mapping from the frequent obejct's original long-form.In this way pragmatics may serve as an essential precursor to more system-level language change.\par
 
 # Acknowledgements
-
-Place acknowledgments (including funding information) in a section at
-the end of the paper.
 

### 2019-01-13 paper/paper.Rmd
commit before sunday additions

@@
 abstract: 
-    Zipf (1935) argued that the relation between a word's frequency and length was 
-    a universal property of human languages emerging from the competing pressures to minimize speaker- and listener-effort during communication.
-    More recent work has explored these ideas, providing additional evidence for a Zipfian analysis of this relation between frequency and length, what Zipf termed
-    the "Law of Abbreviation". However, this work, including Zipf's original analysis, has primarily relied on large-scale observational data, 
-    only providing indirect evidence for the causal role speaker- and listener-pressures have on the "Law of Abbreviation". 
-    In an effort to isolate these competiting dynamics experimentally,
-    Kanwal et al. (2017) conducted a study using an artificial language learning paradigm. The authors showed that 
-    only when both a speaker- and listener-pressures were present did the Law of Abbreviation emerge. The authors
-    highlight that while their data is consistent with a Zipfian analysis, it is also consistent with an alternative, pragmatic language-use account.
-    We consider this latter account, conducting a Bayesian data analysis using the data from Kanwal et al. (2017) representing
-    subjects as rational, pragmatic agents using the Rational Speech Act framework. Fitting subject-level parameters
-    we find good fit between model posterior-predictive values and data from Kanwal et al. (2017). We argue this
-    analysis provides evidence for a connection between Zipfian notions of efficiency driven language change and Gricean 
-    notions of rational-pragmatic langauge use.
+    Zipf (1935) argued that the relation between a word's frequency and length was a universal property of human languages.
+This relation, what Zipf called the “Law of Abbreviation”, emerged from the competing pressures to minimize speaker- and listener-effort during communication.
+Subsequent work has primarily adopted methodological tools in line with Zipf’s original work -- relying on analyses of large-scale observational data.
+Such analyses, however, can only provide indirect evidence for the causal role speaker- and listener-pressures have on the “Law of Abbreviation.”  In an effort to isolate these competing dynamics experimentally, Kanwal et al.
+(2017) conducted a study using an artificial language learning paradigm.
+Only when both a speaker- and listener-pressures were present in a communication game, did the “Law of Abbreviation” emerge.
+Crucially, the authors highlight that their data is consistent with two possible analyses -- a Zipfian account and a pragmatic, language use account -- but that it is impossible to distinguish between the two given nature of the data collected.
+In this work, we consider the pragmatic, language-use account by conducting a Bayesian data analysis, modeling subjects in Kanwal et al.
+(2017) study using the Rational Speech Act framework.
+Fitting subject-level parameters we find good fit between model posterior-predictive values and experimental production data.
+We consider these results in light of arguments put forth in Kanwal et al.
+(2017) with respect to the relation between pragmatic language use and large-scale language change.
 
@@
 
+Kanwal et al. (2017)
+
+Frank & Goodman (2012)
+
+Goodman & Frank (2016)
+
+Horn (1984)
+
+Zipf (1935)
+
+Grice (1957)
+
 ```{r}

### 2019-01-13 paper/paper.Rmd
full RD for review by readers.

@@
 abstract: 
-    Zipf (1935) argued that the relation between a word's frequency and length was a universal property of human languages.
+    Zipf (1935) argued that the relation between a word's frequency and length was a universal property of human language.
 This relation, what Zipf called the “Law of Abbreviation”, emerged from the competing pressures to minimize speaker- and listener-effort during communication.
-Subsequent work has primarily adopted methodological tools in line with Zipf’s original work -- relying on analyses of large-scale observational data.
+Subsequent projects have primarily adopted methodological tools in line with Zipf’s original work -- relying on analyses of large-scale, observational data.
 Such analyses, however, can only provide indirect evidence for the causal role speaker- and listener-pressures have on the “Law of Abbreviation.”  In an effort to isolate these competing dynamics experimentally, Kanwal et al.
@@
 In this work, we consider the pragmatic, language-use account by conducting a Bayesian data analysis, modeling subjects in Kanwal et al.
-(2017) study using the Rational Speech Act framework.
+(2017) using the computational-pragmatics framework.
 Fitting subject-level parameters we find good fit between model posterior-predictive values and experimental production data.
@@
 keywords:
-    "Add your choice of indexing terms or keywords; kindly use a semi-colon; between each term."
+    "Language change; Pragmatics; Bayesian data analysis; Bayesian cognitive modeling"
 
@@
 Zipf (1935) presented a view of human behavior that focused on *effort minimization* as a guiding principle.
-Relying on empirical evidence from corpora, he argued that the unique dynamics of speaker- and listener-effort minimization give rise to the distributional forms found  in natural language.
+Relying on empirical evidence from language corpora, he argued that the unique dynamics of speaker- and listener-effort minimization give rise to the distributional forms found  in natural language.
 Among these, Zipf examined the relation between a word's frequency in a corpus and a variety of properties including its rank-frequency, its denotation size, and its length.
-In terms of this last property, what he called the "Law of Abbreviation," Zipf claimed that the more frequent words are, the shorter they tend to be.
-Evidence for this relationship can be found across languages, may also be present in animal communication systems  (Ferrer-i-Cancho et, 2013), and has also been observed in artificial (programming) languages (CITATION from Kanwal).\par
+In terms of this last property, what he called the "Law of Abbreviation," (LOA) Zipf claimed that the more frequent words are, the shorter they tend to be.
+Evidence for this relationship can be found across languages (CITATION), may also be present in animal communication systems  (Ferrer-i-Cancho et, 2013), and has also been observed in artificial (programming) languages (CITATION from Kanwal).\par
 
@@
 Kanwal et al.
-(2017) highlight the fact that Zipf’s original work, as well as more recent studies analysing large-scale, observational corpora, “do not explicitly target the hypothesized role of communicative pressures…” \par
+(2017) highlight the fact that Zipf’s original work, as well as more recent studies analysing large-scale, observational corpora, “do not explicitly target the hypothesized role of communicative pressures…”  In other words, it is difficult to make causal claims about the role that competing speaker- and listener-pressures play purely from large-scale, observational data analysis.\par
 
-To gain traction on the question of the causal role of speaker-listener pressures, Kanwal et al.
+To gain traction on the causal role of speaker-listener pressures, Kanwal et al.
 (2017) implemented an artificial language learning paradigm to isolate the pressures directly.
-In line with Zipf's original analysis, they hypothesized that only when *both* speaker- and listener- pressures are present in a communication setting would the "Law of Abbreviation" emerge.
+In line with Zipf's original analysis, they hypothesized that only when *both* speaker- and listener- pressures are present in a communication setting would the LOA emerge.
 The authors adopted a miniature artificial language learning set-up in which participants learned a language consisting of three words, which could be used to refer to two items.
@@
 The third word, an abbreviation of the longer forms, could be used to refer to either of the two objects (it was ambiguous).\par
-
-Operationalizing production cost as a function to word length, they hypothesized that when given a choice between alternatives, subjects should prefer to use the shorter, less costly term to refer to a more frequently occurring object.
-Importantly, however, this choice should only appear when there was both a speaker- and listener-pressures present.
+Operationalizing production cost as a function of word length, they hypothesized that when given a choice between alternatives, subjects should prefer to use the shorter, less costly term to refer to a more frequently occurring object.
+Importantly, however, this choice should only appear when there were both speaker- and listener-pressures present.
 Results indicated that speakers did tend to use the less costly form for the more frequent item when both pressures were present, but, importantly, not in conditions where only one or neither of the pressures were present.\par
@@
 (2017) represents an important experimental step in confirming Zipfian notions about lexicon-level efficiency, the authors highlight an important and subtle dimension to these results.
-Zipf’s “Law of Abbreviation” is a lexicon-level theory.
+Zipf’s LOA is a lexicon-level theory.
 That is, it attempts to explain why particular word forms are mapped to particular meanings.
-However, their results are, in theory, consistent with a pragmatic language-use account.
-We refer the reader to their paper for a more detailed discussion, but include the essential elements here:\newline
-
-*"There is a distinction between a language-user's mental representation of the lexicon, and the form-meaning mappings they actually produce in communication...the nature of the communication task in this experiment may have caused them to produce only the short form for one object and the long form for the other based on purely pragmatic considerations."*\newline
-
-In other words, the underlying lexicon learned by subjects may have remained unchanged during the experiment -- a finding which, on its face, is at odds with the claim that this experiment demonstrates lexicon-level change.
+However, their experimental results are, in theory, consistent with a pragmatic language-use account as well.
+That is, the underlying lexicon learned by subjects may have remained unchanged during the experiment -- a finding which, on its face, is at odds with the claim that this experiment demonstrates lexicon-level change.
 Rather, subjects may have *used* the lexicon pragmatically, preferring to map less costly forms to more frequent objects, while leaving the underlying lexicon unchanged.
@@
 
-The current project addresses the question of whether the patterns of production behavior observed in Kanwal et al.
+The current project addresses the question of whether patterns of production behavior observed in Kanwal et al.
 (2017) can be explained by a pragmatic, language-use  account.
@@
 To preview our results we find good fit from model posterior-predictive values to Kanwal et al.
-(2017) participant data ($r^2=0.987$), over and above what would would be expected from a baseline model ($r^2=0.684$).
+(2017) participant data  ($r^2=0.972$), over and above what would would be expected from a baseline model ($r^2=0.466$).
 We believe this provides compelling evidence in support of the alternative hypothesis outlined by Kanwal et al.
@@
 \par
-
 The paper proceeds as follows.
@@
 Following this work we return the question posed by Kanwal et al.
-(2017) of how pragmatic language use might bootstrap language change and offer an argument for how this type of process could in theory unfold.
-We end with a summary of our findings and next steps.\par
+(2017) of how pragmatic language use might bootstrap language change and offer an argument for how this type of process could in theory unfold.\par
 
 ```{r plot-inferred-params, fig.env = "figure*", fig.pos = "h", fig.width=6, fig.height=3, fig.align = "center", set.cap.width=T, num.cols.cap=2, fig.cap = "TEMP - NEEDS UPDATING Inferred subject-level parameter distributions.
-Moving clock-wise from the upper-right facet -- inferred alpha values for subjects, the recursive depth for each subject, the probability that subjects are paying attention during the experiment, the inferred cost / referrent frequency structure used by the subject."}
+Moving clock-wise from the upper-right facet -- inferred alpha values for subjects, the recursive depth for each subject, the probability that subjects are paying attention during the experiment, the inferred cost / referent frequency structure used by the subject."}
 img <- png::readPNG("figs/inferred-params-plot.png")
@@
 
-# Experimental evidence of Zipf's Law of Abbreviation
+# Kanwal et al. (2017) and experimental evidence for Zipf's Law of Abbreviation
 
@@
 In the typical set-up, participants are trained on a simple language -- a series of form-meaning mappings, often using nonce words with novel objects.
-During test, participant’s language producing is recorded, typically over a series of communication events or across multiple generations of participants accomplishing some task (CITATIONS).
-Subsequent analyses focus on how properties of the produced language change as it used over time participants.\par
+During test, participant’s language production is recorded, typically over a series of communication events or across multiple generations of participants accomplishing some task (CITATIONS).
+Subsequent analyses focus on how properties of the produced language change as it is used over time.\par
 
@@
 Participants were assigned to one of four possible conditions which varied in the pressures that were present.
-Both speaker- and listener-pressures were present in the *combined* condition, only the listener pressure present in the *accuracy* condition, only the speaker pressure present in the *time* conding.
+Both speaker- and listener-pressures were present in the *combined* condition, only the listener pressure present in the *accuracy* condition, and only the speaker pressure present in the *time* condition.
 They also tested a fourth *neither* condition in which neither speaker nor listener pressures were present.
-For the purpose of our current project, we do not report our results on the control, *neither* condition as behavior should be arbitrary.\par
+For the purpose of our current project, we do not report our results on the control, *neither* condition, as behavior should be arbitrary.\par
 
 For a comprehensive description of the experimental methodology we refer the reader to Kanwal et al.
-(2017), focusing on essential details in this section.
+(2017), focusing on essential details here.
 During training participants learned labels for two novel objects.
@@
 Crucially, to transmit the label to the matcher the director had to click on it, holding the mouse while each character appeared sequentially.
-In this setting, transmitting a longer-form required more time -- an elegant operationalization of speaker production cost.
+In this setting, transmitting a longer-form required more time (in the *combined* and *time* conditions) -- an elegant operationalization of speaker production cost.
 Listener cost was operationalized via a pressure for the matcher to correctly identify the referent.\par
@@
 In this condition, participants should always prefer to use the unambiguous, long-form labels to maximize matching performance.
-Results across all three conditions were largely consistent with these predictions -- only in the *combined* speaker-listener pressure condition did the Law of Abbreviation emerge via shortening of the label for the most frequent object, but not the infrequent object.
+Results across all three conditions were largely consistent with these predictions, but with clear variance in subject-specific strategies by condition.
+Importantly, only in the *combined* speaker-listener pressure condition did the LOA emerge in many participant’s lexicons via a preference to map the shorter form to the most frequent object, (and not the infrequent object).
 
@@
 
-While these results are consistent with a Zipfian account of the Law of Abbreviation, Kanwal et al.
+While these results are consistent with a Zipfian account of the LOA, Kanwal et al.
 (2017) highlight an alternative explanation.
 They describe a distinction between the "mental representation" of the lexicon and produced forms.
-That is, Zipf's Law of Abbreviation is an account of lexical change, not language-use.
-If participants are retaining both the long-forms and short-form word in their "mental lexicon," but *using the lexicon pragmatically*, then the lexicon itself has not "changed." This analysis falls largely in line with Gricean notion of natural language pragmatics.\par
+In their own words,\newline
+
+*"There is a distinction between a language-user's mental representation of the lexicon, and the form-meaning mappings they actually produce in communication...the nature of the communication task in this experiment may have caused them to produce only the short form for one object and the long form for the other based on purely pragmatic considerations."*\newline
+
+Put differently, Zipf's Law of Abbreviation is an account of language structure, not necessarily language-use.
+If participants are retaining both the long-form and short-form words in their "mental lexicon," but *using the lexicon pragmatically*, then the lexicon itself has not "changed." This analysis falls largely in line with Gricean notion’s of natural language pragmatics.\par
+
+## Grice and rational, pragmatic language-use
 
@@
 They know the short label is faster (less costly to the speaker) than using a long-form, and that the frequent object occurs three times more often than the infrequent object.
-In a given trial, in which the matcher is trying to complete the task a Gricean description of their reasoning might go as follows: \newline
+In a given trial, in which the matcher is trying to complete the task, a Gricean description of their reasoning might go as follows: \newline
 
@@
 
-To model this type of rational, pragmatic interaction we adopt the Rational Speech Act Framework (Frank & Goodman, 2012; Goodman & Frank, 2016).
-RSA is a recursive Bayesian model of pragmatic language production and interpretation, which can largely be seen as a mathematical formalization of essential Gricean principles.
-RSA has proven to be a productive framework for modeling a range of pragmatic phenomena from hyperbole, metaphor, scalar implicature and others (CITATIONS).
-Under RSA a speaker agent describes a conditional distribution, mapping meanings $u\in U$ to utterances $m\in M$, written as $S(u|m)$.
-A speaker agent describes a conditional distribution mapping from utterances to meaning, written as $L(m|u)$.
-Importantly, each of these functions is described in terms of the other.
+To model this type of rational, pragmatic interaction we adopt the Rational Speech Act framework (RSA) (Frank & Goodman, 2012; Goodman & Frank, 2016).
+RSA is a recursive Bayesian model of pragmatic language use, which can largely be seen as a mathematical formalization of essential Gricean principles.
+RSA has proven to be a productive framework for modeling a range of pragmatic phenomena in both language production and language use including hyperbole, metaphor, implicature and others (CITATIONS).
+\par
+
+In the RSA framework, a "speaker agent” defines a conditional distribution, mapping meanings $u\in U$ to utterances $m\in M$, written as $S(u|m)$.
+We consider a prior over utterances $P(U)$ as well as a prior over meanings $P(M)$.
+A ``listener agent” defines a conditional distribution mapping from utterances to meanings, written as $L(m|u)$.
+To capture recursive reasoning between interlocutors, each of these functions is described in terms of the other.
 That is,
@@
 \end{equation}
-Where
+where
 \begin{equation}
@@
 \end{equation}
-And 
+and 
 \begin{equation}
@@
 Defining nested speaker and listener agents could, in principle, lead to infinite regress.
-We take a *literal listener*, denoted $L_0(m|u)$, as a base-case.
+RSA defines a *literal listener*, denoted $L_0(m|u)$, as a base-case.
 The literal listener does not reason about a speaker model, rather this agent considers the literal semantics of the utterance.
@@
 \end{equation}
-Where
+with
 \begin{equation}
@@
 
-## Bayesian data analysis with RSA linking function
+# Bayesian data analysis with RSA linking function
 
 To assess the degree to which the results in Kanwal et al.
-(2017) can be described by a pragmatic, language-use account, we conduct a Bayesian data analysis, using RSA as our linking function.
+(2017) can be described by a pragmatic, language-use account, we conduct a Bayesian data analysis, using RSA to model subject data.
 Our analysis proceeds in two parts.
 In the *parameter inference* phase we infer subject-level parameters.
-In the *model-checking and posterior prediction* phase we assess how well our model, with the inferred parameters, describes the experimental data.\par
+In the *model-checking and posterior prediction* phase we assess how well our model, with the inferred parameters, predicts the actual production data.\par
 
-We model each participant as an RSA agent, inferring for each the following set of parameters -- $\alpha$, $\beta$ are RSA model parameters corresponding the the subjects rationality and recursive depth, respectively.
+## Parameter inference
+We model participants as RSA agents, inferring for each the following set of parameters -- $\alpha$ and $\beta$ are RSA model parameters corresponding the the subjects rationality and recursive depth, respectively.
 The parameters $\lambda$ and $\gamma$ correspond to experiment-specific parameters.
-We describe each parameter in more depth.
+We describe each parameter in more depth here,
 
-#### Rationality parameter ($\alpha$)
-Rational speech-act theory frames natural language understanding and generation as probabilistic inference.
-At the heart of this model is a utility calculation (equation 2) in which the speaker considers both the informativity of using an utterance $u$ to convey some meaning $m$, as well as it’s cost $cost(u)$.
-We capture the degree to which the speaker has as preference to choose the most informativity utterance by multiplying this utility function by a scalar $\alpha$.
+#### Rationality ($\alpha$)
+RSA frames natural language understanding and generation as probabilistic inference.
+At the heart of this model is a utility calculation (equation 2) in which the speaker considers both the informativity of using an utterance $u$ to convey some meaning $m$, as well as its cost ($cost(u)$).
+We capture the degree to which the speaker has as preference to choose the most informative utterance by multiplying the utility function by a scalar $\alpha$.
 We infer a separate $\alpha$ value for each subject using a uniform prior over the integers $1...30$.\par
@@
 Perhaps most characteristic of RSA is the nested model definitions in which speakers and listeners are defined recursively in terms of one another.
-The parameter $\lambda$ controls the depth of recursion.
+The parameter $\beta$ controls the depth of recursion.
 We assume a uniform prior over the integers $1..3$ inferring the level of recursion for each subject.\par
@@
 For each condition we describe the particular utterance cost and meanings prior structure.
-In the *combined* condition we consider utterance costs such that the longer forms are twice as expensive as shorter forms and meaning priors the reflect the training proportions (the infrequent item occurs 8 times and the frequent items 24 times).
-In the *time condition* we assume the same structure except we only consider an $S_0$ speaker who does not consider a listener agent.
-(Recall that no listener agent is present in that condition.) For the *accuracy* condition we consider the same structure as in the *combined* condition, except we assume that there is no cost difference between the utterances.
-We also consider a fourth condition which we call *other*, which allows us to describe distintinctly suboptimal behavior.
+In the *combined-strategy* condition we consider *utterance costs* in which longer forms are twice as expensive as shorter forms and *meaning priors* reflect actual training proportions from Kanwal et al.(2017) (the infrequent item occurs 8 times and the frequent item 24 times).
+In the *time-strategy* condition we assume the same structure as the *combined-strategy* condition except we only consider an $S_0$ speaker who does not consider a listener agent.
+(Recall that no listener agent is present in the experimental *time* condition.) For the *accuracy-strategy* we consider the same structure as in the *combined-strategy* condition, except we assume that there is no cost difference between the utterances.
+We also consider a fourth condition *other-strategy* condition, which allows us to describe distintinctly suboptimal production behavior.
 Some respondents nearly always used the shorter form for the infrequent object.
-This behavior is consistent with reversing the prior object frequencies.
-We use a categorical prior over these cost-meaning structure assuming that the the *other* condition is far less likely than the others.\par
+This behavior is consistent with having reversed the object prior frequencies.
+We use a categorical prior over these cost-meaning structure assuming that the the *other-strategy* condition is far less likely than the others, equivalent to a Dirichlet prior with alpha values $\alpha = [10, 10, 10, 1]$.\par
 
@@
 Finally, we consider that subjects may be employing a rational strategy, but might be distracted during the experiment.
-Our attention parameter $\lambda$ captures the probability that a participants chooses an utterance randomly.
-We use a $\beta(12, 1)$ prior to encode that we expect participants to be paying attention and unlikely to choose randomly during the experiment.\par
+Our attention parameter $\gamma$ captures the probability that a participant chooses an utterance randomly in a given trial.
+We use a beta prior $Beta(a=12, b= 1)$ prior to encode the expectation that participants are likely to pay attention and are unlikely to choose randomly during the experiment.\par
 
-The top-right facet of Figure 1 shows the inferred parameters across all participants.
-In particular, we note that few participants were categorized as using the suboptimal *other* condition strategy and that we inferred that most participants were paying attention throughout the course of the experiment ($\alpha$ levels are well above 0.9).
+### Inferred parameter values
+We approximate the posterior distribution over parameter values with $n=200$ samples for each subject, using the Metropolis-Hasting algorithm (CITATION).
+Figure 1 plots the distribution of inferred parameters across all participants.
+The top-left facet displays the inferred alpha values, which appear largely diffuse across participants.
+The top-right facet plots recursive depth, showing that most participants are identified as having just a single level of recursion.
+The bottom-left facet plots the inferred attention ($\gamma$) indicating that the vast majority of participants were assigned high-attention values.
+The bottom-right facet plots the inferred cost and meaning structures.
+Note that we infer that many participants behaved in a way that was consistent with no cost difference between the short- and long-forms (*accuracy-strategy*) and very few behaved in the strictly sub-optimal *other-strategy*.
+We believe this particular distribution is consistent with the actual variance in strategies employed by participants in Kanwal et al.
+(2017), many of whom displayed production strategies that were not strictly condition-optimal.
 
-```{r plate-notation, fig.env = "figure", fig.pos = "H", fig.align='center', fig.width=2, fig.height=2, set.cap.width=T, num.cols.cap=1, fig.cap = "THIS IS A PLACE-HOLD AND IS INCORRECT"}
+```{r plate-notation, fig.env = "figure", fig.pos = "H", fig.align='center', fig.width=2, fig.height=2, set.cap.width=T, num.cols.cap=1, fig.cap = "THIS IS JUST A PLACE-HOLDER"}
 img <- png::readPNG("figs/plate-temp.png")
@@
 ```
-We infer  each subject we use $n=200$ samples using Metropolis-Hasting to approximate the posterior distribution over parameters.
-Figure 1.
-displays inferred subject params across the all participants.
-
 ### Posterior prediction and model checking
@@
 Intuitively if these predictions do not match the *actual* data used to fit the model, then we have a poor model.
-Recall that in the first step of our Bayesian data analysis we infer parameters for each subject including the rationality parameter ($\alpha$), their recursive depth ($\beta$),  the cost and referent condition structure ($\gamma$), and their level of attention ($\lambda$).\par
+Recall that in the first step of our Bayesian data analysis we infer parameters for each subject including the rationality parameter ($\alpha$), their recursive depth ($\beta$),  the cost and referent condition structure ($\lambda$), and their level of attention ($\gamma$).\par
 
-```{r plot-post-predictives, fig.env = "figure*", fig.pos = "h", fig.width=6, fig.height=3, fig.align = "center", set.cap.width=T, num.cols.cap=2, fig.cap = "Posterior predictive values lead to far better fit to human data than a baseline model.
-Horizontal axes display counts of short-form usage from subjects in Kanwal et al.
+```{r plot-post-predictives, fig.env = "figure*", fig.pos = "h", fig.width=6, fig.height=3, fig.align = "center", set.cap.width=T, num.cols.cap=2, fig.cap = "Posterior predictive values lead to significantly better fit to human data than a baseline model.
+Horizontal axes display proportion of short-form usage from subjects in Kanwal et al.
 (2017).
 Vertical axes display model predictions.
-The left facet displays predicted short-form usage under a baseline model that implements the 'optimal' solution for each condition.
-The right facet displays predicted short-form usage under our BDA model."}
+The left facet displays predicted short-form usage under a baseline model that implements the 'optimal’ Zipfian language for each condition.
+The right facet displays predicted short-form usage under our BDA-RSA model."}
 img <- png::readPNG("figs/bda-vs-baseline-plot.png")
@@
 ```
-Having inferred these subject-level parameters we’d like to ``run” our RSA-participants back through the experiment.
-Intuitively this means that we consider the actual sequence of trials for each subject.
-For a particular trial we examine the particular object they are asked to refer to and pass this two our RSA speaker model with that subject’s inferred parameters, generating a corresponding utterance (i.e.
-we sample $u \sim S_{\text{subject}_i,\text{ trial}_j}(u|m)$).
-In this way we can compare trial-by-trial production data of the RSA agents to actual subject data.
+Having inferred subject-level parameters we would like to “run” our RSA-subjects through the original experiment.
+Operationally, this means that we consider the actual sequence of trials for each human subject, generating a corresponding sequence of utterances for each of the trials they saw.
+For example we consider the $j^{th}$ trial for subject $i$.
+We pass the actual object referent subject $i$ observed in the $jth$ trial ($m_{ij}$) to our RSA instantiation of that subject and generate an utterance ($u_{ij})$.
+This amounts to sampling an utterance $u_{ij} \sim S(u_{ij}|m_{ij}; \theta_i)$ where $\theta_i$ contains the particular parameters inferred for subject $i$ (i.e.
+$\theta_i = \{\alpha_i, \beta_i, \lambda_i, \gamma_i\}$).
+Having “run” each of our RSA subject-representations through the experiment we can compare model production data to actual Kanwal et al.
+(2017) participant production data.
 
@@
 Following the analysis in Kanwal et al.
-(2017) we examine the proportion of trials in which a particular subject used the short-form (see Kanwal Figure 5).
+(2017) we examine the proportion of trials in which a particular subject used the short-form (see Kanwal Figure 3).
 That is, we compare the proportion of trials in which subject $i$ used the short form, comparing this to proportion of trials predicted by our RSA model and by a baseline comparison.
@@
 As a baseline we compare a deterministic speaker using the optimal Zipfian language in each condition.
-Under this model a speaker in the *combined* condition should always map the short form the most frequent item, but not the infrequent item; in the *time* condition the speaker should always use the short form, regardless of the referent; in the *accuracy* condition the speaker should always use the long forms.
-This baseline describes a speaker who deterministically chooses an utterance based on the optimal strategy.
+Under this model, a speaker in the *combined* condition should always map the short form to the most frequent item, but not the infrequent item.
+In the *time* condition, the speaker should always use the short form, regardless of the referent.
+In the *accuracy* condition, the speaker should always use the long form.
+This baseline describes a speaker who deterministically chooses an utterance based on the optimal strategy.\par
 
-To assess model performance we correlate model posterior predictive values for our RSA-subjects and baseline mode to human data.
+To assess model performance we correlate model posterior predictive values for our RSA- and baseline-models to human data.
 Table 2 displays $r^2$ values for both models.
-We find close fit for our rational pragmatic agents, significantly above our baseline model
+We find close fit for our RSA agents, significantly above our baseline model (NEED STAT TEST), accounting for a large proportion of participant-level variance.
+Figure 3 displays model fit between baseline and BDA-RSA predictions to human data.\par
 
@@
 Kanwal et al.
-(2017) presented an experimental test of the Zipfian hypothesis conducting a miniature language-learning experiment and isolating the impact a speaker- and listener-pressures directly.
+(2017) presented an experimental test of the Zipfian hypothesis, conducting a miniature language-learning experiment and isolating the impact of speaker- and listener-pressures directly.
 In their analysis Kanwal et al.
-(2017) highlighted that their results were consistent with a Zipfian, language-change account as well as a pragmatic langauge-use account.
-This distinction, the argued, cannot be assessed with the current set of results as the data collected only constitutes production data, there is not information about the underlying lexicon's the subjects are using.par
+(2017) highlighted that their results were consistent with a Zipfian, language-change account as well as a pragmatic language-use account.
+This distinction, they argued, cannot be assessed with the current set of results as the data collected only included production behavior (and no information about actual, underlying lexicons being used by subjects).\par
 
 To test the pragmatic, language-use account we conducted a Bayesian data analysis, modeling subjects as rational pragmatic agents using the Rational Speech Act framework (Frank & Goodman, 2012; Goodman & Frank, 2016).
-Inferring subject-level parameters we find good fit with the empirical data providing evidence that subject behavior in this experiment may a result of pragmatic langauge use, rather than lexicon change.\par
+Inferring subject-level parameters we find good fit with the empirical data providing evidence that subject behavior in this experiment may be a result of pragmatic language use, rather than lexicon change.\par
 
@@
 Figure 5.
-Of Kanwal et al.
-(2017) shows temporal effects throughout the experiment -- speakers are more likely to use the shorter form later in the experiment compared to earlier in the experiment.
+of Kanwal et al.
+(2017) shows temporal effects throughout the experiment -- speakers are more likely to use the shorter form later in the experiment compared to earlier in the experiment in the *combined* condition.
 Our current model does not capture this particular effect.
-However, there are various ways we might try to instantiate temporal effects corresponding to this kind of convergence.
-For example, this might be instantiated via reduction in uncertainty over the prior frequency distributions for the objects referents.
-That is, uncertainty at the social-epistemic level -- does your partner have the same priors you do.
-Taking another angle we might assume that the cost of the long-form becomes more onerous as the task proceeds, making speakers even more likely to want to avoid it.
-Our current framework provides many opportunities for this type of follow-up and we consider this as an important area for further investigation\par
+However, this effect could be due to a variety of factors modeleable in our current framework.
+One such perspective might relate to uncertainty over object frequency priors diminishing as the experiment progresses.
+In the current framework, a preference to use the short-form in the *combined* condition is modulated by the relative proportion of the frequent and infrequent object -- holding all else constant, the greater the relative difference in object frequencies, the greater the preference to use the short-form.
+If participants were uncertain about the prior frequencies (or that they shared the same priors with their interlocutor) but, became more confident over trials that the more frequent object would appear in the same proportion then we should their preference to use the short-form for that object increase.\par
+A separate approach might consider the temporal effect as emerging from increased sensitivity to production cost as the task proceeds.
+Essentially, waiting for the message to transmit may become more onerous over time, making speakers more likely to want to avoid it.
+Our current framework provides ample opportunity to investigate these and related hypotheses about the exact nature of participant behavior.\par
 
-### The connection between rational langauge use and langauge change
-Recent work has started to draw connection betwen local, in-the-moment language use as modeled by rational, pragmatic agents and aggregate, distribuation-level properties of natural langauge (Goodman & Frank, 2016; Levy, 2018).
+### The connection between rational language use and language change
+Recent theoretical and computational work has highlighted connections between local, in-the-moment language use as modeled by rational, pragmatic agents and aggregate, distribution-level properties of natural language (Goodman & Frank, 2016; Levy, 2018).
 This is not without precedent.
@@
 He describes a set-up in which there are a set of utterance $U$ and meanings $M$ which differ in costs and frequency, respectively.
-While Zipf was not more explicitly in diescribing the level at which language optimization occurs, his examples mirror the reference game setting we have adopted here.
-Morever, the explicit connection with rational, pragmatic langauge use was described by Horn (1984) who argued that Zipf's speaker and listener pressures ("Speaker's economy" and "Auditor's ecnonomy") were fundamental to Grice's (1957) derivation of his Cooperative Principle and associated conversational maxims.
-In general, it is not hard to see a scenario in which pragmatic language use bootstraps later lexicon-level change via processes like inter-generational transmission.
+While Zipf did not explicitly described the level at which language optimization occurs, his example mirrors the reference game setting we have adopted here.\par
+
+We note that the connection between Zipfian principles and Gricean, conversational pragmatics was first described by Horn (1984), who argued that Zipf's speaker and listener pressures ("Speaker's economy" and "Auditor's economy") were fundamental to Grice's (1957) derivation of his Cooperative Principle and associated conversational maxims.
+In general, it is not hard to imagine a scenario in which pragmatic language use bootstraps later lexicon-level change via processes like intergenerational transmission.
 Kanwal et al.
-(2017) present this very idea writing,"pragmatics-driven asymmetry in usage may or may not lead to an immediate shift in lexical representations, it may be an important first step in such a change." Take the exmaple of the current experiment.
-In the *combined* some subjects produced langauges which only consisted of the shortened form for the frequent object and the long form for the infrequent object.
-If a subsequent generation of speakers were to learn from this data, they would never be exposed to the original mapping from the frequent obejct's original long-form.In this way pragmatics may serve as an essential precursor to more system-level language change.\par
+(2017) present this very idea writing,"[whether] pragmatics-driven asymmetry in usage may or may not lead to an immediate shift in lexical representations, it may be an important first step in such a change." Take the example of the current experiment.
+In the *combined* condition some subjects produced languages which only consisted of the shortened form for the frequent object and the long form for the infrequent object.
+If a subsequent generation of speakers were to learn from this data, they would never be exposed to the original mapping from the frequent object's original long-form.
+In this way pragmatics may serve as an essential precursor to more system-level language change.\par
 
 # Acknowledgements
+Fill in
 
 # References
-
-Kanwal et al. (2017)
-
-Frank & Goodman (2012)
-
-Goodman & Frank (2016)
-
-Horn (1984)
-
-Zipf (1935)
-
-Grice (1957)
 

### 2019-01-23 bda/analysis/fit_model.Rmd
separate model fitting and plotting

--- 2019-01-08 bda/analysis/fit_model.Rmd
@@
 
-```{r libraries}
+```{r libraries, warning=FALSE, message=FALSE}
 library(ggplot2)
@@
 
-```{r}
+```{r helpers}
 # getModelFile()
@@
 
-```{r}
-# All data
-f_all_data <- "../data/counts_all.csv"
-dfAll <- read.csv(f_all_data)
-cat("n=", length(dfAll$IP))
-```
-
-```{r}
-# f_bda_model <- "../models/bda.wppl"
-# model_bda <- getModelFile(f_bda_model)
-```
-
-```{r}
+```{r track_problem_ips}
+# Problem IPs saving just to track
 # problemIPs <- c('142.196.192.44', 
@@
 
-```{r}
-# pt <- proc.time()
-# df_bda <- rwebppl::webppl(model_bda, data=dfAll, data_var='rData', cores=4)
-# et <- proc.time() - pt
-# cat(paste("Run time: ", et))
+```{r model-run-data-loading}
+RUN_MODEL <- FALSE
 
-df_predictives <- read.csv("../analysis/df_predictives_20181119.csv")
-```
+# User set this
+# -------------
+MODEL_TYPE <- "NAIVE_BASELINE"
+# -------------
 
-```{r merge bda and raw data}
-df_predictives <- merge(dfAll, df_predictives)
-# write_csv(df_predictives, 'df_predictives_20181119.csv')
-```
+# Experimental data we'd like to model
+f_kanwal_counts_data <- "../data/counts_all.csv"
+# Note this data is restricted to "accuracy", "combined" and "time" sections.
+dfAll <- read.csv(f_kanwal_counts_data)
+cat("n=", length(dfAll$IP), "\n")
 
-Check out Model / Human `r^2`
-```{r}
-df_frequent <- df_predictives %>% 
-  select(IP, condition, alpha, depth, inferredCondition, frequent_cnt, posteriorPredictiveFrequent) %>%
-  mutate(type='frequent') %>%
-  rename(actualCnt=frequent_cnt,
-         estCnt=posteriorPredictiveFrequent)
-
-df_infrequent <- df_predictives %>% 
-  select(IP, condition, alpha, depth, inferredCondition, infrequent_cnt, posteriorPredictiveInfrequent) %>%
-  mutate(type='infrequent') %>%
-  rename(actualCnt=infrequent_cnt,
-         estCnt=posteriorPredictiveInfrequent)
-
-df_predictives2 <- rbind(df_frequent, df_infrequent)
-cor(df_predictives2$actualCnt, df_predictives2$estCnt)^2
-
+if (RUN_MODEL) {
+  pt <- proc.time()
+  # Baseline or Full BDA
+  model_file <- getModelFileFromType(MODEL_TYPE)
+  model_bda <- getModelFile(model_file)
+  df_bda <- rwebppl::webppl(model_bda, data=dfAll, data_var='rData', cores=4)
+  # Write file
+  write.csv(df_bda, file=file.path("../analysis", paste0("df_", MODEL_TYPE, ".csv")))
+  et <- proc.time() - pt
+  cat(paste("Run time: ", et))
+}
 ```
@@
 ```{r}
-plot_all_post_preds <- df_predictives2 %>%
-  # mutate(actualFrequentProps=frequent_cnt / frequent_n,
-  #        actualInrequentProps=infrequent_cnt / infrequent_n,
-  #        estFrequentProps=posteriorPr5edictiveFrequent / frequent_n,
-  #        estInfrequentProps=posteriorPredictiveInfrequent / infrequent_n) %>%
-  # select(IP, condition, inferredCondition, actualFrequentProps, estFrequentProps, actualInrequentProps, estInfrequentProps) %>%
-  # gather(type, value, c(posteriorPredictiveFrequent, posteriorPredictiveInfrequent))
-  ggplot(aes(x=actualCnt, y=estCnt, col=condition), alpha=0.4) +
-    geom_smooth(aes(group=0), method='lm') +
-    geom_point(aes(fill=condition), alpha=0.6, size=3, colour="white", pch=21) +
-    # geom_point(aes(x=actualInrequentProps, y=estInfrequentProps, col=condition), alpha=0.4) +
-    ylab('Posterior predictive \n counts of short form') +
-    xlab('Human data \n counts of short form') +
-    ggtitle('r^2=0.987') +
-    theme_few() +
-    theme(plot.title = element_text(hjust = 0.5))
-
-ggsave('../../paper/figs/all-post-predictives-plot.png', plot_all_post_preds, width=6, height=3)
-```
-
-```{r}
-df_predictives2 %>%
-  # mutate(actualFrequentProps=frequent_cnt / frequent_n,
-  #        actualInrequentProps=infrequent_cnt / infrequent_n,
-  #        estFrequentProps=posteriorPr5edictiveFrequent / frequent_n,
-  #        estInfrequentProps=posteriorPredictiveInfrequent / infrequent_n) %>%
-  # select(IP, condition, inferredCondition, actualFrequentProps, estFrequentProps, actualInrequentProps, estInfrequentProps) %>%
-  # gather(type, value, c(posteriorPredictiveFrequent, posteriorPredictiveInfrequent))
-  ggplot(aes(x=actualCnt, y=estCnt, col=condition), alpha=0.4) +
-    geom_smooth(method='lm') +
-    geom_point(aes(fill=condition), alpha=0.6, size=3, colour="white", pch=21) +
-    # geom_point(aes(x=actualInrequentProps, y=estInfrequentProps, col=condition), alpha=0.4) +
-    facet_wrap(~condition, scales="free") +
-    ylab('Posterior predictive \n counts of short form') +
-    xlab('Human data \n counts of short form') +
-    theme_few()
-
-ggsave('../../paper/figs/post-predictives-facet-plot.png', width=6, height=3)
-```
-
-Compare to "optimal baseline" optimal in that `time` condition always pick short, `accuracy` always long, etc.
-```{r}
-bda_vs_baseline_plot <- df_predictives2 %>%
-  mutate(baseline=ifelse(condition=='combined' & type=='frequent', 24,
-                         ifelse(condition=='combined' & type=='infrequenty', 8,
-                                ifelse(condition=='accuracy' & type=='frequent', 0,
-                                       ifelse(condition=='accuracy' & type=='infrequent', 24,
-                                              ifelse(condition=='time' & type=='frequent', 48, 16)))))) %>%
-  gather('model', 'pred', c(baseline, estCnt)) %>%
-  mutate(model=ifelse(model=='estCnt', 'BDA', 'Baseline')) %>%
-  ggplot(aes(x=actualCnt, y=pred, col=condition), alpha=0.4) +
-    geom_point(aes(fill=condition), alpha=0.6, size=3, colour="white", pch=21) +
-    geom_abline(intercept=0, slope=1, col='grey', alpha=0.4) +
-    # geom_point(aes(x=actualInrequentProps, y=estInfrequentProps, col=condition), alpha=0.4) +
-    facet_wrap(~model, scales="free") +
-    ylab('Baseline model \n counts of short form') +
-    xlab('Human data \n counts of short form') +
-    theme_few()
-
-ggsave('../../paper/figs/bda-vs-baseline-plot.png', bda_vs_baseline_plot, width=6, height=3)
-```
-
-```{r}
-df_predictives2 %>%
-  mutate(baseline=ifelse(condition=='combined' & type=='frequent', 24,
-                         ifelse(condition=='combined' & type=='infrequenty', 8,
-                                ifelse(condition=='accuracy' & type=='frequent', 0,
-                                       ifelse(condition=='accuracy' & type=='infrequent', 24,
-                                              ifelse(condition=='time' & type=='frequent', 48, 16)))))) %>%
-  group_by(condition) %>%
-  summarise(r2 = cor(actualCnt, estCnt)^2)
-
-df_predictives2 %>%
-  mutate(baseline=ifelse(condition=='combined' & type=='frequent', 24,
-                         ifelse(condition=='combined' & type=='infrequenty', 8,
-                                ifelse(condition=='accuracy' & type=='frequent', 0,
-                                       ifelse(condition=='accuracy' & type=='infrequent', 24,
-                                              ifelse(condition=='time' & type=='frequent', 48, 16)))))) %>%
-  summarise(r2 = cor(actualCnt, baseline))
-
-```
-
-## Subject-level inferred params?
-
-```{r}
-plot_alpha <- df_predictives %>%
-  ggplot(aes(alpha)) +
-    geom_density() +
-    theme_few()
-
-plot_depth <- df_predictives %>%
-  ggplot(aes(depth)) +
-    geom_histogram(bins=3) +
-    theme_few()
-
-plot_attention <- df_predictives %>%
-  ggplot(aes(inferredAttention)) +
-    geom_density() +
-    theme_few()
-
-plot_condition <- df_predictives %>%
-  mutate(inferredCondition=as.character(inferredCondition)) %>%
-  mutate(inferredCondition=ifelse(inferredCondition=='conventionalized', 'other', inferredCondition)) %>%
-  group_by(inferredCondition) %>%
-  summarise(count=n()) %>%
-  ggplot(aes(x=inferredCondition, y=count)) +
-    geom_bar(stat='identity', position='dodge') +
-    theme_few() +
-    theme(axis.text.x = element_text(angle = 45, hjust = 1))
-
-plot_all_params <- grid.arrange(plot_alpha, plot_depth, plot_attention, plot_condition, nrow=2)
-ggsave('../../paper/figs/inferred-params-plot.png', plot=plot_all_params, width=6, height=4)
-# plot_condition <- df_predictives %>%
-#   group_by(inferredCondition) %>%
-#   summarise(n=n())
-```
-
-How well do we infer the strategies?
-```{r}
-mean(df_predictives$inferredCondition == df_predictives$actualCondition)
-```
-
-We use conventionatlized only for a few. These are very strange in that they *almost* always avoid using the short form for the frequent term, but use it for the non-frequent 
-```{r}
-df_predictives %>%
-  filter(inferredCondition=='conventionalized')
-```
-
-Output inferred subject-level params
-```{r}
-df_predictives %>%
-  select(IP, alpha, depth, inferredCondition) %>%
-  toJSON() %>%
-  write_lines('subjectInferredParams.json')
-```
-
-# Try to model time dynamics
-
-```{r}
-f_temporal_bda_model <- "../models/bdaFig5.wppl"
-model_temporal_bda <- getModelFile(f_temporal_bda_model)
-df_trajectories <- rwebppl::webppl(model_temporal_bda)
-```
-
-```{r}
-
-df2 <- data.frame()
-for (i in seq(1, length(df_trajectories))) {
-  df2 <- rbind(df2, df_trajectories[[i]])
+# data2predictions()
+# ==================
+# Minor pre-processing for correlation analysis.
+#
+# Parameters
+# ----------
+# modelFile: str
+#  Model type.
+# 
+# Returns
+# -------
+# file.path
+#    File path to model.
+#
+data2predictions <- function(modelType) {
+  df <- read.csv(getModelFittedModelData(modelType))
+  df_frequent <- df %>%
+    select(IP, actualCondition, actualFrequent, posteriorPredictiveFrequent, inferredAttention) %>%
+    mutate(type='frequent') %>%
+    rename(actualCnt=actualFrequent,
+           estCnt=posteriorPredictiveFrequent)
+  df_infrequent <- df %>%
+    select(IP, actualCondition, actualInfrequent, posteriorPredictiveInfrequent, inferredAttention) %>%
+    mutate(type='infrequent') %>%
+    rename(actualCnt=actualInfrequent,
+           estCnt=posteriorPredictiveInfrequent)
+  df_combined <- rbind(df_frequent, df_infrequent) %>%
+    mutate(modelType=modelType)
+  df_combined
 }
 
+# deterministicBaselineDf()
+# =========================
+# Kanwal et al optimal deterministic strategy.
+#
+# Parameters
+# ----------
+# df: Data.Frame
+#   Data frame of model predictions
+# Returns
+# -------
+# Data.Frame
+#    Data frame with deterministic condition-optimal data.
+#
+deterministicBaselineDf <- function(df) {
+  df %>%
+    select(IP, actualCondition, actualCnt, type) %>%
+    mutate(estCnt=ifelse(actualCondition=='combined' & type=='frequent', 24,
+                         # Only use short form for frequent in combined
+                         ifelse(actualCondition=='combined' & type=='infrequent', 0,
+                                # Never use short form during accuracy
+                                ifelse(actualCondition=='accuracy' & type=='frequent', 0,
+                                       ifelse(actualCondition=='accuracy' & type=='infrequent', 0,
+                                              # Always use short form during time
+                                              ifelse(actualCondition=='time' & type=='frequent', 48, 16)))))) %>%
+    mutate(modelType='DETERMINISTIC BASELINE',
+           inferredAttention=1)
+}
 ```
-```{r}
-numIPs <- length(unique(df2$IP))
-df2$trialNum <- rep(seq(1, 31), numIPs)
-```
-
-```{r}
-df2 %>%
-  filter(trialNum==1 | trialNum==31) %>%
-  spread(trialNum, data) %>%
-  rename('first'=`1`, 'last'=`31`) %>%
-  mutate(diff=last-first) %>%
-  ggplot(aes(x=IP, y=diff)) +
-    geom_bar(stat='identity', position='dodge') +
-    theme_few()
-```
-
-```{r}
-df2 %>%
-  group_by(trialNum) %>%
-  summarise(n=n(),
-            avgProp=mean(data),
-            ci_high=avgProp + 1.96 *sqrt(avgProp*(1-avgProp)/n),
-            ci_low=avgProp - 1.96 *sqrt(avgProp*(1-avgProp)/n)) %>%
-  ggplot(aes(x=trialNum, y=avgProp)) +
-    geom_line() +
-    geom_errorbar(aes(ymax=ci_high, ymin=ci_low)) +
-    ylim(0, 1) +
-    theme_few()
-```

### 2019-01-23 bda/analysis/generate_plots.Rmd
separate model fitting and plotting

---
title: "Paper_plots"
author: "Ben"
date: "1/22/2019"
output: pdf_document
---

```{r libraries, warning=FALSE, message=FALSE}
library(ggplot2)
library(ggthemes)
library(gridExtra)
library(rwebppl)
library(tidyverse)
```

```{r helpers}
# getModgetModelFittedModelData()
# ===============================
# Get cached data from model fitting.
#
# Parameters
# ----------
# modelFile: str
#  Model type.
# 
# Returns
# -------
# file.path
#    File path to model.
#
getModelFittedModelData <- function(typ) {
  dirPath <- '../data/model_predictions_data'
  file.path(dirPath, paste0("df_", typ, ".csv"))
}
```

# Analysis
```{r combine-all-models}
df_naive <- data2predictions("NAIVE_BASELINE")
df_deterministic_baseline <- deterministicBaselineDf(df_naive)
df_no_recursion <- data2predictions("NO_RECURSION")
# df_full <- data2predictions("FULL_MODEL")

df_all_models <- rbind(df_naive, df_deterministic_baseline, df_no_recursion)
```

All conditions plot
```{r}
df_all_models %>% 
    rename(`Object referrent`=type,
           `Experimental condition`=actualCondition) %>%
    mutate(modelType=ifelse(modelType=="DETERMINISTIC BASELINE", "Deterministic baseline", 
                            ifelse(modelType=="NAIVE_BASELINE", "Zipf language +\n attention", "RSA-BDA"))) %>%
    mutate(modelType=factor(modelType, levels=c("Deterministic baseline", "Zipf language +\n attention", "RSA-BDA"))) %>%
    ggplot(aes(x=actualCnt, y=estCnt, col=`Experimental condition`), alpha=0.4) +
      geom_point(aes(fill=`Experimental condition`, shape=`Object referrent`), alpha=0.6, size=3) +
      geom_abline(intercept=0, slope=1, col='grey', alpha=0.4) +
      ylab('Model predicted \n counts of short form') +
      xlab('Human data \ncounts of short form') +
      theme_few() +
      facet_grid(~modelType)
```

Combined condition
```{r}
df_all_models %>% 
  filter(actualCondition=="combined") %>%
  ggplot(aes(x=actualCnt, y=estCnt, col=actualCondition), alpha=0.4) +
    geom_point(aes(fill=actualCondition, shape=type), alpha=0.6, size=3) +
    geom_abline(intercept=0, slope=1, col='grey', alpha=0.4) +
    ylab('Baseline model \n counts of short form') +
    xlab('Human data \n counts of short form') +
    theme_few() +
    facet_grid(~modelType)
```

Correlations
```{r}
df_all_models %>% 
  group_by(modelType) %>%
  summarise(r2=cor(estCnt, actualCnt)^2)

df_all_models %>% 
  group_by(modelType, actualCondition) %>%
  summarise(r2=cor(estCnt, actualCnt)^2)
```

```{r}
# 20190122 BP lesion test
df_frequent <- df_bda %>%
  select(IP, actualCondition, actualFrequent, posteriorPredictiveFrequent, inferredAttention) %>%
  mutate(type='frequent') %>%
  rename(actualCnt=actualFrequent,
         estCnt=posteriorPredictiveFrequent)

df_infrequent <- df_bda %>% 
  select(IP, actualCondition, actualInfrequent, posteriorPredictiveInfrequent, inferredAttention) %>%
  mutate(type='infrequent') %>%
  rename(actualCnt=actualInfrequent,
         estCnt=posteriorPredictiveInfrequent)

df_predictives2 <- rbind(df_frequent, df_infrequent)
cor(df_predictives2$actualCnt, df_predictives2$estCnt)^2
```

## Subject-level inferred params?

#### alpha plot
```{r}
df_alphas <- rbind(df_naive, df_no_recursion)

df_alphas %>%
  ggplot(aes(inferredAttention, col=modelType, fill=modelType)) +
    geom_density(alpha=0.3) +
    theme_few()

```

#### all params
```{r}
df_RSA <- read.csv(getModelFittedModelData("NO_RECURSION"))

plot_alpha <- df_RSA %>%
  group_by(alpha) %>%
  summarise(n=n()) %>%
  ggplot(aes(x=alpha, y=n)) +
    geom_bar(stat='identity') +
    xlab("Alpha") +
    ylab("Count") +
    theme_few()

plot_attention <- df_RSA %>%
  ggplot(aes(inferredAttention)) +
    geom_density(alpha=0.3, fill='blue') +
    xlab("Attention") +
    ylab("Density") +
    theme_few()

plot_condition <- df_RSA %>%
  mutate(inferredCondition=as.character(inferredCondition)) %>%
  mutate(inferredCondition=ifelse(inferredCondition=='conventionalized', 'other', inferredCondition)) %>%
  group_by(inferredCondition) %>%
  summarise(count=n()) %>%
  ggplot(aes(x=inferredCondition, y=count)) +
    geom_bar(stat='identity', position='dodge') +
    xlab("Condition") +
    ylab("Count") +
    theme_few() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))

bottom_row <- plot_grid(plot_condition, plot_alpha, labels = c('B', 'C'), align = 'h', rel_widths = c(1, 1.3))
plot_all_params <- plot_grid(plot_attention, bottom_row, labels = c('A', ''), ncol = 1, rel_heights = c(1, 1.2))
# ggsave('../../paper/figs/inferred-params-plot.png', plot=plot_all_params, width=6, height=4)
```

# Save below
-------------

### Proportions Analysis

```{r}
df_cors <- df_predictives2 %>%
  mutate(baseline=ifelse(condition=='combined' & type=='frequent', 24,
                         ifelse(condition=='combined' & type=='infrequent', 0,
                                ifelse(condition=='accuracy' & type=='frequent', 0,
                                       ifelse(condition=='accuracy' & type=='infrequent', 0,
                                              ifelse(condition=='time' & type=='frequent', 48, 16)))))) %>%
  gather('model', 'pred', c(baseline, estCnt)) %>%
  group_by(IP, condition, actualCnt, model) %>%
  summarise(cntShortForms=sum(pred)) %>%
  group_by(IP, model, condition) %>%
  summarise(estCnts=sum(cntShortForms),
            actCnts=sum(actualCnt)) %>%
  mutate(condition_n=ifelse(condition=='combined' | condition=='accuracy', 32, 64)) %>%
  mutate(estProp=estCnts/condition_n,
         actProp=actCnts/condition_n) %>%
  ungroup
```

```{r}
df_cors %>%
  group_by(model) %>%
  summarise(n=n(),
            r=cor(estProp, actProp)^2)
```

## Proportions plot1

```{r}
df_cors %>%
  mutate(model=ifelse(model=='baseline', 'Baseline', 'BDA RSA')) %>%
  ggplot(aes(x=actProp, y=estProp, col=condition), alpha=0.4) +
    geom_point(aes(fill=condition), alpha=0.6, size=3) +
    geom_abline(intercept=0, slope=1, col='grey', alpha=0.4) +
    # geom_point(aes(x=actualInrequentProps, y=estInfrequentProps, col=condition), alpha=0.4) +
    facet_wrap(~model) +
    ylab('Baseline model \n counts of short form') +
    xlab('Human data \n counts of short form') +
    ylim(0, 1) +
    xlim(0, 1) +
    theme_few()

ggsave('../../paper/figs/bda-vs-baseline-plot.png', plot = last_plot(), width=6, height=3)
```

Some kind of significance test?

```{r}
fisher.z <- function(r1,r2,n1,n2) {
  # Todo: review (https://stats.stackexchange.com/questions/64152/are-two-pearson-correlation-coefficients-different)
  ((0.5*log((1+r1)/(1-r1)))-(0.5*log((1+r2)/(1-r2))))/((1/(n1-3))+(1/(n2-3)))^0.5
}

```

```{r}
rs <- df_predictives2 %>%
  mutate(baseline=ifelse(condition=='combined' & type=='frequent', 24,
                         ifelse(condition=='combined' & type=='infrequent', 8,
                                ifelse(condition=='accuracy' & type=='frequent', 0,
                                       ifelse(condition=='accuracy' & type=='infrequent', 24,
                                              ifelse(condition=='time' & type=='frequent', 48, 16)))))) %>%
  filter(type=='frequent') %>%
  gather(model, value,  c(estCnt, baseline)) %>%
  select(actualCnt, model, value, condition) %>%
  group_by(model) %>%
  summarise(n=n(),
            r=cor(actualCnt, value)^2)
print(rs)
```

How well do we infer the strategies?
```{r}
mean(df_predictives$inferredCondition == df_predictives$actualCondition)
```

We use conventionatlized only for a few. These are very strange in that they *almost* always avoid using the short form for the frequent term, but use it for the non-frequent 
```{r}
df_predictives %>%
  filter(inferredCondition=='conventionalized')
```

Output inferred subject-level params
```{r}
df_predictives %>%
  select(IP, alpha, depth, inferredCondition) %>%
  toJSON() %>%
  write_lines('subjectInferredParams.json')
```

# Try to model time dynamics

```{r}
f_temporal_bda_model <- "../models/bdaFig5.wppl"
model_temporal_bda <- getModelFile(f_temporal_bda_model)
df_trajectories <- rwebppl::webppl(model_temporal_bda)
```

```{r}

df2 <- data.frame()
for (i in seq(1, length(df_trajectories))) {
  df2 <- rbind(df2, df_trajectories[[i]])
}

```
```{r}
numIPs <- length(unique(df2$IP))
df2$trialNum <- rep(seq(1, 31), numIPs)
```

```{r}
df2 %>%
  filter(trialNum==1 | trialNum==31) %>%
  spread(trialNum, data) %>%
  rename('first'=`1`, 'last'=`31`) %>%
  mutate(diff=last-first) %>%
  ggplot(aes(x=IP, y=diff)) +
    geom_bar(stat='identity', position='dodge') +
    theme_few()
```

```{r}
df2 %>%
  group_by(trialNum) %>%
  summarise(n=n(),
            avgProp=mean(data),
            ci_high=avgProp + 1.96 *sqrt(avgProp*(1-avgProp)/n),
            ci_low=avgProp - 1.96 *sqrt(avgProp*(1-avgProp)/n)) %>%
  ggplot(aes(x=trialNum, y=avgProp)) +
    geom_line() +
    geom_errorbar(aes(ymax=ci_high, ymin=ci_low)) +
    ylim(0, 1) +
    theme_few()
```

### 2019-01-23 bda/analysis/fit_model.Rmd
beginning RD re-rog

@@
 ```{r model-run-data-loading}
-RUN_MODEL <- FALSE
+RUN_MODEL <- TRUE
 
@@
 }
+
+# Problems
+# 108.196.136.202 - worked once
+# 117.229.19.52 - worked once
+# 122.161.26.31
+# 140.141.208.86
+# 80.101.212.2
+# 24.113.49.186
 ```

### 2019-01-23 paper/paper.Rmd
update to new cogsci styleing
beginning RD re-rog

@@
 Among these, Zipf examined the relation between a word's frequency in a corpus and a variety of properties including its rank-frequency, its denotation size, and its length.
-In terms of this last property, what he called the "Law of Abbreviation," (LOA) Zipf claimed that the more frequent words are, the shorter they tend to be.
-Evidence for this relationship can be found across languages (CITATION), may also be present in animal communication systems  (Ferrer-i-Cancho et, 2013), and has also been observed in artificial (programming) languages (CITATION from Kanwal).\par
+In terms of this last property, what he called the "Law of Abbreviation," (LoA) Zipf claimed that the more frequent words are, the shorter they tend to be.\par
 
-Zipf argued that this relationship between frequency and length was an emergent property of competing speaker-listener pressures.
-Under this framing, communication is viewed as a cooperative, joint activity between interlocutors.
-However, this joint activity contains a fundamental asymmetry -- what is effortful for a speaker (production) is different from what is effortful for a listener (understanding).
-An optimal speaker language, Zipf argued, would consist of a single, low-cost word which was fully ambiguous; referring to all possible objects in the world.
-Adopting such a language, a listener would need to disambiguate every utterance by the speaker.
-By contrast, if a language were optimised solely in terms of listener effort, the language should bijectively map all forms to meanings so there was no need for a listener to disambiguate.\par
-
-Zipf argued that these competing forces, what he called “Speaker’s economy” and “Auditor’s economy”, give rise to the particular distributional forms found in all languages, including the relationship between word frequency and length.
-While this framework is compelling and can largely be seen as foundational to a family of theories examining efficiency in language-structure (CITATIONS) and  -use (CITATIONS), the issue is far from settled.
-Subsequent work has argued that properties such as the relation between word frequency and length can theoretically be explained by a process of random typing (Ferrer-i-Cancho & Mascoso del Pardo, 2012).
+Zipf argued that this relationship between frequency and length was an emergent property of competing speaker-listener pressures what he called “Speaker’s economy” and “Auditor’s economy”.
+Subsequent work has argued that the LoA and related properties can theoretically be explained by a process of random typing (Ferrer-i-Cancho & Mascoso del Pardo, 2012).
 Kanwal et al.
-(2017) highlight the fact that Zipf’s original work, as well as more recent studies analysing large-scale, observational corpora, “do not explicitly target the hypothesized role of communicative pressures…”  In other words, it is difficult to make causal claims about the role that competing speaker- and listener-pressures play purely from large-scale, observational data analysis.\par
+(2017) highlight that Zipf’s original work, as well as more recent studies analysing large-scale, observational corpora, “do not explicitly target the hypothesized role of communicative pressures…” In other words, it is difficult to make causal claims about the role that competing speaker-listener pressures play from observational data analysis.
 
@@
 (2017) implemented an artificial language learning paradigm to isolate the pressures directly.
-In line with Zipf's original analysis, they hypothesized that only when *both* speaker- and listener- pressures are present in a communication setting would the LOA emerge.
-The authors adopted a miniature artificial language learning set-up in which participants learned a language consisting of three words, which could be used to refer to two items.
-Crucially, two of the words were longer (seven versus three characters) than a third word.
-The two longer words could only be used to refer to one of the two objects.
-The third word, an abbreviation of the longer forms, could be used to refer to either of the two objects (it was ambiguous).\par
-Operationalizing production cost as a function of word length, they hypothesized that when given a choice between alternatives, subjects should prefer to use the shorter, less costly term to refer to a more frequently occurring object.
-Importantly, however, this choice should only appear when there were both speaker- and listener-pressures present.
-Results indicated that speakers did tend to use the less costly form for the more frequent item when both pressures were present, but, importantly, not in conditions where only one or neither of the pressures were present.\par
+In line with Zipf's original analysis, they hypothesized that only when *both* speaker- and listener- pressures are present in a communication setting would the LoA emerge.
+The authors adopted a miniature artificial language learning paradigm in which participants learned a language and completed a reference game task.
+The experiment included three conditions which directly manipulated speaker- and listener- pressures.
+They found that the language produced in the conditions were largely consistent with a Zipfian account -- only in the condition in which both speaker- and listener-pressures were present did production data follow LoA predictions.\par
 
 While Kanwal et al.
-(2017) represents an important experimental step in confirming Zipfian notions about lexicon-level efficiency, the authors highlight an important and subtle dimension to these results.
-Zipf’s LOA is a lexicon-level theory.
-That is, it attempts to explain why particular word forms are mapped to particular meanings.
-However, their experimental results are, in theory, consistent with a pragmatic language-use account as well.
-That is, the underlying lexicon learned by subjects may have remained unchanged during the experiment -- a finding which, on its face, is at odds with the claim that this experiment demonstrates lexicon-level change.
-Rather, subjects may have *used* the lexicon pragmatically, preferring to map less costly forms to more frequent objects, while leaving the underlying lexicon unchanged.
-The authors’ highlight that because they only recorded participants’ actual language production they cannot distinguish between the lexicon-change account and pragmatic, language-use account.\par
+(2017) results are consistent with a Zipfian account of language change they highlight that their results are also consistent with a pragmatic, language-use account.
+That is, the results may reflect pragmatic use of an \textit{unchanging lexicon}.
+This observation is consistent with recent work which has begun to probe the question of how pragmatic processes, occuring in-the-moment of communication may give rise to long term effects in language -- perhaps such as those described by Zipf.
+Goodman & Frank (2016) highlight that understanding in an individual communication event can alter learning processes and lead to systematic changes in word meanings (Smith et al., 2013).
+Further, words can be pragmatically extended or narrowed in their denotations.
+The effect of such pragmatic use over time may give rise to longer-time lexical-level change.
 
-The current project addresses the question of whether patterns of production behavior observed in Kanwal et al.
-(2017) can be explained by a pragmatic, language-use  account.
-To do so, we conduct a Bayesian data analysis, modeling subjects as rational pragmatic agents using the Rational Speech-Act framework (RSA) (Frank & Goodman, 2017; Goodman & Frank, 2016).
-To preview our results we find good fit from model posterior-predictive values to Kanwal et al.
-(2017) participant data  ($r^2=0.972$), over and above what would would be expected from a baseline model ($r^2=0.466$).
-We believe this provides compelling evidence in support of the alternative hypothesis outlined by Kanwal et al.
-(2017) -- that subjects may have been behaving pragmatically, rather than demonstrating real change at the level of language-structure.
+Kanwal et al.
+(2017) give us an opportunity to explore the question of how long-term lexical change might be related to in-the-moment pragmatics.
+As pointed out by the authors, the experimental results are, in theory, consistent with both a Zipfian language-change account as well as pragmatic language-use account.
+Because they only recorded participants’ actual language production behavior, however, they cannot distinguish between these accounts using normal data-analytic methods.
+To do so would require access to the underlying \textit{mental representation} of the lexicon used by participants.
+The Rational Speech Act framework (RSA) (Frank & Goodman, 2012; Goodman & Frank, 2016) provides a useful computational formalism to gain traction on the question of pragmatics posed by Kanwal et al.
+(2017).
+RSA can largely be understood as a formulation of Gricean pragmatics, casting natural language understanding and production as probabilistic inference in agents who reason about one-another.
 \par
+
+The current project compares a rational, pragmatic account as instantiated by RSA and a Zipfian, language-change account in which probabilistic agents use a modified lexicon with some error-probability.
+We conduct a Bayesian data analysis, inferring subject-level parameters and comparing posterior-predictive values for each participant’s data.
+To preview our results we find good fit from the pragmatics account, both overall ($r^2=XXX$), and in the crucial speaker- and listener-effort condition ($r^2=XXXX$); above what we find from the “Zipfian, language change” model ($r^2=XXXXX$).
+
+We interpret these results in two ways -- first we believe it supports the importance of pragmatics in the particular study design of Kanwal et al.
+(2017).
+More generally, however, and in line with the authors’ discussion, we believe these results engage deeper questions on the relationship between pragmatics and language-change, in which pragmatica language use can be seen as an important first step in bootstrapping later change.\par
+
 The paper proceeds as follows.
@@
 (2017), their hypotheses and the alternative interpretation of their results.
-We then introduce our general approach to the Bayesian data analysis as well as the Rational Speech-act framework we will use to model subject’s pragmatic language use.
+We then introduce our general approach to the Bayesian data analysis as well as the computational instantiations of the two theories with discussion of the Rational Speech-act framework.
 We evaluate our results summarising the findings and highlighting dimensions of their analysis our current implementation does not yet capture.
